### 2023-03-20 analysis/spatiotemporal-transmission-risk.Rmd
data, workflow and minor fixes

---
title: "Spatiotemporal transmission risk"
author: "Nicolas Banholzer"
date: "2022-11-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
```

## Libraries

```{r}
library(tidyverse)
library(reshape2)
library(rriskDistributions)
library(tidybayes)
library(lubridate)
library(sf)
library(raster)
library(terra)
library(ggridges)
library(grid)
library(ggplot2)
library(ggrepel)

source("../utils/distr.r")
source("../utils/trans_risk.r")
source("../utils/spatial.r")
source("../utils/plotting.R")

# plot with and height of input and data graphs
width_descr <- 10.5 / cm(1)
height_descr <- 7.5 / cm(1)
text_descr <- 12
```

## Data

```{r}
selected_dates <- list.files("../simulations/Masi/")
selected_dates <- as.Date(selected_dates[!grepl("old", selected_dates)])
```

```{r}
# overview
files_by_date <- data.frame(
  directory = list.files("../data-clean/Masi/combined-data", full.names = T)
) %>%
  mutate(date = basename(directory),
         files_in_dir = map(directory, list.files),
         Clinical = map_lgl(files_in_dir, function(x) any(grepl("clinical", x))),
         CO2 = map_lgl(files_in_dir, function(x) any(grepl("environmental", x))),
         Tracking = map_lgl(files_in_dir, function(x) any(grepl("tracking", x))))

files_by_date %>%
  dplyr::select(date, Clinical, CO2, Tracking) %>%
  reshape2::melt("date") %>%
  ggplot(aes(x = variable, y = date, fill = value)) +
  geom_tile() +
  labs(x = "Dataset", fill = "Available") +
  scale_fill_manual(values = wes_palette("Chevalier1")[c(2,1)]) +
  theme_classic() +
  theme(axis.title.y = element_blank())
ggsave(filename = "../insights/data-check-dataset-availability.pdf", width = 16 / cm(1), height = 21 / cm(1))
```

### Clinical

```{r}
clinical_dta <- readRDS("../data-clean/Masi/clinical-data/clinical-data.rds") %>%
  filter(date %in% selected_dates)

# number of registered patients
clinical_dta %>%
  group_by(patient_id) %>%
  slice(1) %>%
  nrow()

# share of TB patients
share_tb <- clinical_dta  %>%
  group_by(date, patient_id) %>%
  summarize(tb = any(tb_suspect == "yes")) %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(n = n(),
            n_tb = sum(tb)) %>%
  arrange(date) %>%
  mutate(prop_tb = n_tb / n * 100,
         date = format(date, format = "%b %d"),
         date = factor(date, levels = unique(date))) %>%
  ungroup() 
share_tb_pl <- share_tb %>%
  ggplot(aes(x = date, y = n_tb)) +
  geom_bar(stat = "identity", fill = "grey", color = "black", width = .5) +
  annotate("text", x = "Nov 12", y = mean(share_tb$n_tb), label = paste0("Mean = ",round(mean(share_tb$n_tb), 1)), color = "darkred", size = 8 / cm(1), vjust = -1) +
  geom_hline(aes(yintercept = mean(share_tb$n_tb)), color = "darkred") +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  scale_x_discrete(breaks = format(selected_dates, format = "%b %d")[c(1, 4, 8, 12)]) +
  labs(y = "Number of masked TB patients") +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        plot.margin = margin(5.5, 5.5, 5.5, 5.5),
        text = element_text(size = text_descr))
share_tb_pl
ggsave(plot = share_tb_pl, filename = "../doc/IWHOD/presentation/figures/share-tb-patients.png", w = width_descr, h = height_descr)
```

### CO2

```{r}
co2_pl <- readRDS("../data-clean/Masi/environmental-data/co2-temp-humid.rds") %>%
  filter(location %in% c("waiting room", "tb room"),
         date %in% selected_dates) %>%
  mutate(location = factor(ifelse(location == "tb room", "TB room", "Waiting room"), levels = c("Waiting room", "TB room")),
         time = as.POSIXct(time, format = "%H:%M", tz = "CET")) %>%
  ggplot(aes(x = time, y = co2, color = location, fill = location)) +
  stat_lineribbon(alpha = .25, size = .75) +
  scale_x_datetime(expand = c(0,0), date_breaks = "2 hours", date_labels = "%H:%M") +
  scale_y_continuous(expand = c(0,0), breaks = scales::breaks_width(width = 100)) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_color_manual(values = wes_palette("Moonrise2")) +
  labs(y = expression("CO"[2]*" (ppm)"), fill = "Quantile", color = "Quantile") +
  theme_classic() +
  theme(legend.position = c(.65,.95), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        plot.margin = margin(5.5, 5.5, 5.5, 5.5), axis.title.x = element_blank(), legend.title = element_blank(),
        text = element_text(size = text_descr)) 
co2_pl
ggsave(plot = co2_pl, filename = "../doc/IWHOD/presentation/figures/co2-levels.png", width = width_descr, height = height_descr)
```

### Building

```{r}
# background image
clinic_img <- png::readPNG("../data-raw/Masi/building/clinic_clipped.png")

# building
cellSize <- sqrt(1 * 1e9 / 3e3) 
building <- terra::vect(paste("../data-raw", "Masi", "building", "clinic-vector.gpkg", sep = "/"))
building_sf <- sf::st_as_sf(building, crs = NA)
sf::st_crs(building_sf) <- NA
waiting_room <- shapeToSpatial(building_sf$geometry[2], cellSize)
waiting_room_df <- fortify(waiting_room)
passage <- shapeToSpatial(building_sf$geometry[4], cellSize)
passage_df <- fortify(passage)
tb_room <- shapeToSpatial(building_sf$geometry[3], cellSize)
tb_room_df <- fortify(tb_room)

# test
# waiting_room_df$value <- rnorm(nrow(waiting_room_df))
# passage_df$value <- rnorm(nrow(passage_df))
# tb_room_df$value <- rnorm(nrow(tb_room_df))
# 
# ggplot(mapping = aes(x = long, y = lat, group = group, fill = value)) +
#   geom_polygon(data = waiting_room_df) +
#   geom_polygon(data = passage_df) +
#   geom_polygon(data = tb_room_df) +
#   ggpubr::background_image(clinic_img) +
#   annotate("text", x = -13200, y = -5300, label = "Entrance", size = 8 / cm(1), fontface = 2, hjust=0,vjust=0) +
#   annotate("text", x = -13400, y = -2300, label = "Waiting room", size = 8 / cm(1), fontface = 2, hjust=0,vjust=0) +
#   annotate("text", x = -5000, y = -4100, label = "Registration", size = 8 / cm(1), fontface = 2, angle = 90, hjust=0,vjust=0) +
#   annotate("text", x = 2600, y = -2700, label = "TB room", size = 8 / cm(1), fontface = 2) +
#   scale_y_continuous(expand = expansion(add=c(2500, 4000))) +
#   scale_x_continuous(expand = expansion(add=c(400, 500))) +
#   labs(fill = "Density") +
#   theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
#         legend.position = c(0.825,0.075), legend.direction = "horizontal", legend.title = element_blank(), #legend.title = element_text(vjust = .75),
#         panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 3, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank())
```

### No. people

```{r}
no_people <- do.call(rbind, lapply(selected_dates, function(x) readRDS(paste0("../data-clean/Masi/combined-data/", x, "/tracking-linked-clinical-data.rds")))) %>%
  mutate(date = as.Date(format(time, format = "%Y-%m-%d")),
         time_min = as.POSIXct(format(time, format = "%H:%M"), format = "%H:%M"))

# number of clinical attendees
no_people %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  nrow()

# over time
no_people_time <- no_people %>%
  group_by(date, time_min, time) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(date, time_min) %>%
  summarize(n = mean(n)) %>%
  ungroup()

no_people_time_pl <- no_people_time %>%
  ggplot(aes(x = time_min, y = n)) +
  stat_lineribbon(alpha = .25, size = .5) +
  scale_x_datetime(expand = c(0,0), date_breaks = "2 hours", date_labels = "%H:%M") +
  scale_y_continuous(expand = c(0,0), limits = c(0, NA), breaks = scales::breaks_width(5)) +
  scale_fill_brewer(palette = "OrRd") +
  labs(y = "No. of clinical attendees", fill = "Quantile", color = "Quantile") +
  theme_bw2() +
  theme(axis.title.x = element_blank(), legend.position = c(.8,.9), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  guides(fill = guide_legend(title.position = "top"))
no_people_time_pl

# Spatial
no_people_spatial_pl <- no_people %>%
  ggplot(aes(x = x, y = y)) +
  geom_bin2d(binwidth = 100) +
  scale_fill_stepsn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"), trans = "log", breaks = c(1, 10, 100, 1000, 10000), labels = function(x) signif(x / nrow(no_people_time) * 60 * 24, 1)) +
    ggpubr::background_image(clinic_img) +
    annotate("text", x = -13200, y = -5800, label = "Entrance", size = text_descr / cm(1), hjust=0,vjust=0) +
    annotate("text", x = -14000, y = -2600, label = "Waiting room", size = text_descr / cm(1), hjust=0,vjust=0) +
    annotate("text", x = -5800, y = -4500, label = "Registration", size = text_descr / cm(1), angle = 90, hjust=0,vjust=0) +
    annotate("text", x = 3000, y = -2700, label = "TB room", size = text_descr / cm(1)) +
    scale_y_continuous(expand = expansion(mult=c(0.05, 0.215))) +
    scale_x_continuous(expand = expansion(mult=c(0.03, 0))) +
    labs(fill = "Density (#tracks/day)") +
    theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.3, "cm"),
          legend.position = c(0.825,0.075), legend.direction = "horizontal",legend.title = element_text(vjust = .75),
          panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(title.position = "top", label.position = "bottom"))
no_people_spatial_pl
ggsave(plot=no_people_spatial_pl, filename = "../doc/IWHOD/presentation/figures/no-people-spatial.png", width = width_descr, height = height_descr)
```

## Assumptions

```{r}
n_samples <- 1000
set.seed(12345)
```

### Spatial emission

* Wang (2021): 
  * Travel distance of particles depends on particle size, activity, and environmental factors
  * Smaller aerosols can travel further, stay in the air for a longer time, and typically contain higher viral load
  * The concentration of exhaled aerosols is highest close to the source and decreases with distance
  * Short-range aerosol transmission difficult to prevent with ventilation
* Chen (2020): Short-range aerosols dominate over most distance up to 2m, droplets only play a role at very close distance <0.5
* Ko (2004): exponential decrease of potential infections on an airplance (high ventilation) with distance to infector

--> Assume an exponential decrease in viral concentration with distance to source relative to the source (relative concentration = 1).

```{r}
relQC_pl <- expand.grid(distance = seq(0, 10, .1), 
                           mu = rMuIc(n_samples)) %>%
  mutate(relQC = dIc(distance, mu) * 100) %>%
  ggplot(aes(x = distance, y = relQC)) +
  stat_lineribbon(alpha = .5, size = .5) +
  scale_x_continuous(expand = c(0,0), limits = c(0, 10), breaks = seq(0, 10, 2)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  scale_fill_brewer(palette = "OrRd") +
  labs(x = "Distance (m)", y = "Rel. prob. (%)", fill = "Quantile") +
  theme_classic() +
  theme(legend.position = c(.9,.85), legend.direction = "vertical", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        legend.spacing.y = unit(.1, "cm"),
        plot.margin = margin(5.5, 10, 5.5, 5.5),
        text = element_text(size = text_descr),
        legend.title = element_blank()) +
  guides(fill = guide_legend(title.position = "top", byrow = T))
ggsave(plot = relQC_pl, filename = "../doc/IWHOD/presentation/figures/spatial-relative-quanta-concentration.png", width = width_descr * 0.66, height = height_descr * 0.66)
```

### Infectious quanta

* Riley (1962): 130 patients, q: 1.25
* Escombe (2008): 117 patients, q: 8.2; but high values above 100 are possible
* Nardell (1991) : 1 patients, q: 12.5
* Andrews (2014) : 571 patients, q: 0.89

--> Take the mean and standard deviation of these values and use a student-t distribution with small degrees of freedom to consider outliers

```{r}
q <- c(1.25, 8.2, 12.5, 0.89)
n_q <- c(130, 117, 1, 571)
mean_q <- sum(n_q * q) / sum(n_q)
sd_q <- sqrt( sum( (n_q / sum(n_q) * (q - mean_q) ^ 2 ) ) )

qLiterature <- data.frame(q = q, 
                          y = .01, 
                          lab = c("Riley", "Escombe", "Nardell", "Andrews"))

q_pl <- data.frame(q = seq(0, 40, .1)) %>%
  mutate(density = dq(q)) %>%
  ggplot(aes(x = q, y = density)) +
  geom_line() +
  geom_vline(mapping = aes(xintercept = 2), linetype = "dashed", color = "darkred") +
  annotate("text", x = 3, y = 0.175, hjust = 0, label = "Mean = 2.0", size = 8 / cm(1), color = "darkred") +
  geom_point(data = qLiterature, mapping = aes(x = q, y = y, color = lab, shape = lab)) +
  geom_text_repel(data = qLiterature, mapping = aes(x = q, y = y, color = lab, label = lab), size = 8 / cm(1)) +
  annotate("text", x = 30, y = 0.175, label = expression("q ~ Student-t("*mu*"=2.0,"*sigma*"=2.5"*","*nu*"=1)"), size = 8 / cm(1)) +
  scale_x_continuous(expand = c(0,0), limits = c(0, 40)) +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  labs(x = "Quanta emission (log q/h)", y = "Density") +
  theme_bw2() +
  theme(plot.margin = margin(5.5, 10, 5.5, 5.5), legend.position = "none")
save_plot(q_pl, pdf_file = "../results/inputs/quanta-emission.pdf", w = aw, h = ah)
```

### Unmasked TB patients

* Berhanu (2023): 
  * study group: clinic attendees living with HIV, contact with TB patients in past year, or had TB in the past 2 years
  * finding: 55% of  with a sputum result positive for M. tuberculosis did not have a positive TB symptom screen
* Hamada (2018): The WHO 4-question symptom screen (cough, fever, weight loss, and night sweats) misses up to half the TB cases among people living with human immunodeficiency virus (HIV) on antiretroviral therapy (ART)
--> Unmasked TB patients: Poisson(mu = No. of TB suspects)

```{r}
unmaskedTB <- data.frame(n_tb = share_tb$n_tb) %>%
  mutate(unmasked_tb = lapply(n_tb, function(x) rTBunmasked(n = n_samples, lambda = x))) %>%
  unnest(unmasked_tb) %>%
  group_by(unmasked_tb) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop_unmasked_tb = n / sum(n)) 
mean_unmaskedTB <- sum(unmaskedTB$unmasked_tb * unmaskedTB$prop_unmasked_tb)
unmaskedTB_pl <- unmaskedTB %>%
  ggplot(aes(y = prop_unmasked_tb, x = unmasked_tb)) +
  geom_bar(stat="identity", fill = "grey", color = "black") +
  geom_vline(mapping = aes(xintercept = mean_unmaskedTB), linetype = "dashed", color = "darkred") +
  annotate("text", y = 0.25, x = mean_unmaskedTB, hjust = -.2, label = paste0("Mean = ", round(mean_unmaskedTB, 1)), size = 8 / cm(1), color = "darkred") +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  scale_x_continuous(expand = c(0,0)) +
  labs(x = "No. of unmasked TB patients", y = "Probability (%)") +
  theme_bw2()
save_plot(unmaskedTB_pl, pdf_file = "../results/inputs/unmasked-tb-patients.pdf", w = aw, h = ah)
```

### Ventilation

Average volume of exhaled gas:
* Richardson et al. (2014): 0.1 - 0.17 l/s per person based on Feher (2017)
--> Use Uniform(0.1, 0.17)

CO2 levels in outdoor air:
* https://gml.noaa.gov/dv/data/index.php?site=CPT&parameter_name=Carbon%2BDioxide,%20accessed%2012%20November%202019
--> Set to 4125.5

CO2 levels in exhaled air:
* Rudnick and Milton (2003): 40,000
--> Set to 40,000

CO2 generation rate
* Persily et al.: see Table 5, around 0.003 to 0.005
--> Use Uniform(0.003, 0.005)

### Inactivation

* Nardell (2016) reviews Loudon et al. (1969) and Lever et al. (2000) reporting contradictory findings of half-lives between 20m to 6h. Exact viability in the air is unknown.
* The finding by Lever et al. would suggest that sustained contact is required for transmission.
* The viability in the air probably depends a lot on the conditions, in particular ventilation and UV.

--> TODO: Check how high the ventilation rate is compared to half-lives as to whether we can ignore the half-lives.

### Deposition onto surfaces

* Wang et al. (2021): The larger the size of the particle, the less long it will reside in the air. Environmental factors such as temperature, humidity and ventilation also play an important role.

## Evaluation

### Transmission risk

```{r}
# transmission risk 
sim_files <- sapply("../simulations/Masi", list.files, full.names = T) 
sim_files <- unlist(sapply(sim_files, list.files, full.names = T))
tr_files <- sim_files[grepl("TR", sim_files)]

average_transRisk <- function(file) {
  # quanta concentration and transmission risk
  qctr <- readRDS(file)

  # individual patient transmission risk by simulation
  tr <- qctr %>%
    group_by(patient_id, date, sim) %>%
    slice(1) %>%
    ungroup()

  return(tr)
}

transRisk <- parallel::mclapply(tr_files, average_transRisk, mc.cores = 6)
transRisk <- do.call(rbind, transRisk)

transRisk_av <- transRisk %>%
    group_by(date, sim) %>%
    summarize(across(c(P_N, P_N_ns), mean)) %>%
    ungroup()

transRisk_av_daily <- transRisk_av %>%
  group_by(date) %>%
  summarize(across(c(P_N, P_N_ns), list(m = mean, s = sd))) %>%
  ungroup()

mean_risk_daily <- round(mean(transRisk_av_daily$P_N_m * 100), 2)
range_daily <- round(range(transRisk_av_daily$P_N_s * 100), 2)
message(sprintf("Spatiotemporal transmission risk across clinical attendee: %s (range: %s percent to %s percent)", mean_risk_daily, range_daily[1], range_daily[2]))

# high risk patients
transRisk_high <- transRisk %>%
  mutate(high_risk = ifelse(P_N > 0.01, 1, 0),
         date = as.Date(date)) %>%
  group_by(date, sim) %>%
  summarize(n_risk = sum(high_risk)) %>%
  ungroup() %>%
  group_by(sim) %>%
  arrange(desc(date)) %>%
  mutate(date_letter = factor(LETTERS[2:(n()+1)], levels = LETTERS[1:(length(selected_dates)+1)])) %>%
  ungroup()

transRisk_high_daily <- transRisk_high %>%
  group_by(date_letter) %>%
  summarize(n_risk_m = mean(n_risk),
            n_risk_s = sd(n_risk),
            n_risk_l = quantile(n_risk, 0.025),
            n_risk_u = quantile(n_risk, 0.975)) %>%
  ungroup()

transRisk_high_daily_av <- data.frame(
  n_risk_m = mean(transRisk_high_daily$n_risk_m),
  n_risk_s = sd(transRisk_high_daily$n_risk_m)) %>%
  mutate(n_risk_u = n_risk_m + qnorm(.975) * n_risk_s,
         n_risk_l = n_risk_m + qnorm(.025) * n_risk_s,
         date_letter = factor(LETTERS[1], levels = LETTERS[1:(length(selected_dates)+1)]),
         n_risk_l = ifelse(n_risk_l < 0, 0, n_risk_l))

ylabs <- c("Average", rev(format(as.Date(selected_dates), format = "%b %d")))

highRiskPat_pl <- ggplot() +
  geom_vline(aes(xintercept = mean(transRisk_high_daily$n_risk_m)), linetype = "dotted", color = "darkred") +
  stat_pointinterval(data = transRisk_high, mapping = aes(x = n_risk, y = date_letter, slab_color = after_stat(level)), 
                    point_interval = "mean_qi", .width = c(.5, .9), point_size = 2, fill = "white", shape = 21, point_color = "darkred") +
  geom_errorbar(data = transRisk_high_daily_av, mapping = aes(xmin = n_risk_l, xmax = n_risk_u, y = date_letter),
                width = .2, color = "darkred") +
  geom_point(data = transRisk_high_daily_av, mapping = aes(x = n_risk_m, y = date_letter),
             shape = 23, color = "darkred", fill = "white", size = 4) +
  scale_color_manual(values = scales::brewer_pal()(3)[-1], aesthetics = "slab_color") +
  scale_x_continuous(expand = c(0.01,0)) +
  scale_y_discrete(breaks = LETTERS[1:(nrow(transRisk_high_daily)+1)], labels = ylabs) +
  labs(x = "Number of attendees") +
  theme_classic() +
  theme(text = element_text(size = text_descr),
        axis.title.y = element_blank(), 
        legend.position = "none")

tab_data <- rbind(transRisk_high_daily, transRisk_high_daily_av) %>%
  mutate(meanCI = paste0(add_ws(round(n_risk_m, 0)), " [", add_ws(round(n_risk_l, 0), n = 1), ", ", add_ws(round(n_risk_u, 0)), "]"))
tab_base <- ggplot(data = tab_data, aes(y = date_letter)) +
    ylab(NULL) + xlab("  ") + 
    theme(plot.title = element_text(hjust = 0, face = 2), ## centering title on text
        axis.text.x=element_text(color="white"), ## need text to be printed so it stays aligned with figure but white so it's invisible
        axis.line=element_blank(),
        axis.text.y=element_blank(),axis.ticks=element_blank(),
        axis.title.y=element_blank(),legend.position="none",
        panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),plot.background=element_blank())
summary_tab <- tab_base +
  geom_text(aes(x = 0, label = meanCI), size = text_descr / cm(1)) +
  ggtitle("Mean (95%-UI)")

highRiskPat_pl_wTable <- grid.arrange(highRiskPat_pl, summary_tab, ncol = 2, heights = height_descr, widths = width_descr * c(0.7,.3))
ggsave(plot = highRiskPat_pl_wTable, filename = "../doc/IWHOD/presentation/figures/high-risk-patients.png", width = width_descr * 1.5, height = height_descr * 1.5)
```

### Quanta concentration

```{r}
# files
qc_files <- sim_files[grepl("QC", sim_files)]

average_quantaConc <- function(file) {
  # quanta concentration and transmission risk
  qctr <- readRDS(file)

  # average quanta concentration
  quantaDF_av <- qctr %>%
    mutate(daytime = ifelse(format(time, "%H:%M") <= "12:00", "Morning", "Afternoon"),
           daytime = factor(daytime, levels = c("Morning", "Afternoon"))) %>%
    group_by(date, daytime, id, location) %>%
    summarize(N = mean(N, na.rm = T)) %>%
    ungroup()

  return(quantaDF_av)
}

quantaConc <- parallel::mclapply(qc_files, average_quantaConc, mc.cores = 6)
quantaConc <- do.call(rbind, quantaConc)

quantaConc_av <- quantaConc %>%
  group_by(date, daytime, location, id) %>%
  summarize(N = mean(N, na.rm = T)) %>%
  ungroup() %>%
  group_by(daytime, location, id) %>%
  summarize(N = mean(N, na.rm = T)) %>%
  ungroup()

quantaConc_av_wr <- quantaConc_av %>%
  mutate(id = as.character(id)) %>%
  dplyr::filter(location == "waiting room",
                id %in% unique(waiting_room_df$id)) 
quantaConc_av_wr <- left_join(waiting_room_df, quantaConc_av_wr, by = c("id"))
n_pa_ids <- max(as.numeric(passage_df$id))
quantaConc_av_pa <- quantaConc_av %>%
  group_by(daytime) %>%
  arrange(id) %>%
  slice((n()-n_pa_ids+1):n()) %>%
  mutate(id = as.character(1:n())) %>%
  ungroup() 
quantaConc_av_pa <- left_join(passage_df, quantaConc_av_pa, by = c("id"))
quantaConc_av_tb <- quantaConc_av %>%
  filter(location == "tb room") %>%
  group_by(daytime) %>%
  mutate(id = as.character(1:n())) %>%
  ungroup() 
quantaConc_av_tb <- left_join(tb_room_df, quantaConc_av_tb, by = c("id"))

entrance_lab <- textGrob(label = "Entrance", x = 0.1, y = 0.165, gp = gpar(size = text_descr / cm(1)), just = c("left"))
wr_lab <- textGrob(label = "Waiting room", x = 0.05, y = 0.45, gp = gpar(size = text_descr / cm(1)), just = c("left"))
tb_lab <- textGrob(label = "TB room", x = 0.75, y = 0.4, gp = gpar(size = text_descr / cm(1)), just = c("left"))
re_lab <- textGrob(label = "Registration", x = 0.44, y = 0.25, gp = gpar(size = text_descr / cm(1)), just = c("left"), rot = 90)

qc_pl <- ggplot(mapping = aes(x = long, y = lat, group = group, fill = N)) +
  geom_polygon(data = quantaConc_av_wr) +
  geom_polygon(data = quantaConc_av_pa) +
  geom_polygon(data = quantaConc_av_tb) +
  ggpubr::background_image(clinic_img) +
  scale_fill_stepsn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"), trans = "log1p", labels = function(x) x * 1000, breaks = c(.1, .5, 1, .15, 2)/1000) +
  scale_y_continuous(expand = expansion(add=c(2500, 3250))) +
  scale_x_continuous(expand = expansion(add=c(400, 500))) +
  facet_wrap(~daytime, ncol = 1) +
  annotation_custom(entrance_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(wr_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(tb_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(re_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  labs(fill = expression("Density (10"^-3*"quanta/m"^3*")")) +
  theme_classic() +
  theme(text = element_text(size = text_descr),
        axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = c(0.825,0.05), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(title.position = "top"))
qc_pl
ggsave(plot = qc_pl, filename = "../doc/IWHOD/presentation/figures/quanta-concentration.png", width = width_descr * 1.5, height = height_descr * 2)
```