### 2023-08-10 illustrations/tracking-examples.Rmd
update tracking plots

---
title: "Tracking Examples"
author: "Nicolas Banholzer"
date: "2023-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(grid)
source("../utils/plotting.r")
```

## Data

```{r}
# building
clinic_img <- png::readPNG("../data-raw/Massi/building/clinic_clipped.png")
building <- terra::vect(paste0("../data-raw/", "Massi", "/building/clinic-vector.gpkg"))
building_sf <- sf::st_as_sf(building, crs = NA)
st_crs(building_sf) <- NA
building_df <- fortify(building_sf)

# files
files <- list.files("../data-clean/Massi/2021/patient-tracking", full.names = T)
files <- files[grepl("rds", files)]

# load data
df <- do.call(rbind, lapply(files, function(x) {
  d <- readRDS(x)
  if ("date_time" %in% colnames(d) & "time" %in% colnames(d)) {
    d <- d %>%
      dplyr::select(-time) %>%
      dplyr::select(patient_id, obs_id_new, obs_id, tracking_end, match_type, date_time, everything())
  } else if ("time" %in% colnames(d)) {
    d <- dplyr::rename(d, date_time = time)
  }
  return(d)
})) %>%
  mutate(patient_id = factor(patient_id),
         date = as.Date(date_time))

# subset of tracks
smpl <- df %>%
  filter(tracking_end == "Entered + Exited") %>%
  group_by(patient_id, date) %>%
  filter(n_distinct(obs_id_new) == 1) %>%
  summarize(duration = as.numeric(difftime(last(date_time), first(date_time)))) %>%
  ungroup() %>%
  filter(duration > 5) %>%
  mutate(selected = T) %>%
  filter(patient_id %in% c(8139, 1433, 339, 510, 1003))

df_smpl <- df %>%
  left_join(smpl) %>%
  filter(selected)

# start and end times of subset
df_smpl_se <- df_smpl %>%
  group_by(patient_id) %>%
  slice(1, n()) %>%
  mutate(type = factor(c("Start", "End"), levels=c("Start", "End"))) %>%
  ungroup()
```

## Plot

```{r}
building_pl <- ggplot(building_df) +
  geom_sf(linewidth = .5, alpha = 0, lwd = 0) +
  scale_fill_discrete(na.translate = F) +
  scale_x_continuous(breaks = seq(-20000, 20000, 1000)) +
  scale_y_continuous(breaks = seq(-20000, 20000, 1000)) +
  guides(fill = "none")

text_descr <- 10
en_lab <- textGrob(label = "Entrance", x = 0.1, y = 0.165, gp = gpar(size = text_descr / cm(1)), just = c("left"))
wr_lab <- textGrob(label = "Waiting room", x = 0.05, y = 0.45, gp = gpar(size = text_descr / cm(1)), just = c("left"))
tb_lab <- textGrob(label = "TB room", x = 0.75, y = 0.4, gp = gpar(size = text_descr / cm(1)), just = c("left"))
re_lab <- textGrob(label = "Registration", x = 0.44, y = 0.25, gp = gpar(size = text_descr / cm(1)), just = c("left"), rot = 90)

examples_pl <- building_pl +
  geom_path(data = df_smpl, mapping = aes(x = x, y = y, group = factor(patient_id), color = factor(patient_id)), alpha = .5) +
  geom_point(data = df_smpl_se, mapping = aes(x = x, y = y, group = factor(patient_id), color = factor(patient_id), shape = type)) +
  ggpubr::background_image(clinic_img) +
  #scale_y_continuous(expand = expansion(add=c(2000, 3250))) +
  #scale_x_continuous(expand = expansion(add=c(800, 0))) +
  scale_shape_manual(values = c(1,13)) +
  annotation_custom(en_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(wr_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(tb_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(re_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  theme_classic() +
  theme(text = element_text(size = text_descr),
        axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.3, "cm"),
        legend.position = c(0.825,0.075), legend.direction = "horizontal",
        panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), legend.title = element_blank())

examples_pl
```