### 2023-12-06 analysis/clinical-data.Rmd
complete preprocessing, first descriptive look into data, updating modeling assumptions

---
title: "Clinical Data"
author: "Nicolas Banholzer"
date: "2023-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)

source("../utils/plotting.r")
```

## Data

```{r}
df <- readRDS("../data-clean/clinical/clinical-data.rds") %>%
  filter(between(month(date), 10, 11))

modeled_days <- list.files("../data-clean/patient-tracking/linked-and-imputed")
modeled_days <- gsub("[[:punct:]]rds", "", modeled_days)
modeled_days <- as.Date(modeled_days)
```

## Descriptives

### Diagnosed TB over time

```{r}
diag_tb <- df %>%
  mutate(`Date modeled` = ifelse(date %in% modeled_days, "Yes", "No")) %>%
  group_by(date, `Date modeled`) %>%
  summarize(n_infectious = sum(is_infectious)) %>%
  ungroup()

diag_tb %>%
  ggplot(aes(x = date, y = n_infectious, fill = `Date modeled`)) +
  geom_bar(stat = "identity") +
  scale_x_date(date_labels = "%b %d", breaks = modeled_days) +
  scale_y_continuous(expand = c(0,0)) +
  theme_custom() +
  theme(legend.position = "top")
```

```{r}
diag_tb %>%
  ggplot(aes(x = n_infectious)) +
  geom_histogram()

#' Assumption:
#' Number of diagnosed ~ number of diagnosed TB
#' --> Assume a Dirichlet prior with support according to the counts in the range (1,5).
```

### Suspected TB over time

```{r}
df %>%
  mutate(`Date modeled` = ifelse(date %in% modeled_days, "Yes", "No")) %>%
  group_by(date, `Date modeled`) %>%
  summarize(n_infectious = sum(was_suspected & !is_infectious)) %>%
  ungroup() %>%
  ggplot(aes(x = date, y = n_infectious, fill = `Date modeled`)) +
  geom_bar(stat = "identity") +
  scale_x_date(date_labels = "%b %d", breaks = modeled_days) +
  scale_y_continuous(expand = c(0,0)) +
  theme_custom() +
  theme(legend.position = "top")
```