### 2024-01-19 analysis/clinical-data.Rmd
revised model and new simulation function

---
title: "Clinical Data"
author: "Nicolas Banholzer"
date: "2023-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)

source("../utils/plotting.r")
```

## Data

```{r}
df <- readRDS("../data-clean/clinical/clinical-data.rds") %>%
  filter(between(month(date), 10, 11))

modeled_days <- list.files("../data-clean/patient-tracking/linked-and-imputed")
modeled_days <- gsub("[[:punct:]]rds", "", modeled_days)
modeled_days <- as.Date(modeled_days)
modeled_days_abbrv <- format(modeled_days, format = "%b %d")
```

## Descriptives

### Diagnosed/Suspected TB over time

```{r}
diag_tb <- df %>%
  mutate(date_in_model = format(as.Date(date), format = "%b %d"),
         date_in_model = ifelse(date_in_model %in% modeled_days_abbrv, date_in_model, "Not modeled"),
         date_in_model = factor(date_in_model, levels = c(modeled_days_abbrv, "Not modeled"))) %>%
  group_by(date, date_in_model) %>%
  summarize(n_infectious = sum(is_infectious),
            n_suspected = sum(was_suspected & !is_infectious)) %>%
  ungroup() %>%
  reshape2::melt(c("date", "date_in_model")) %>%
  mutate(variable = ifelse(grepl("infectious", variable), "Diagnosed", "Suspected"),
         variable = factor(variable, levels = c("Suspected", "Diagnosed")))

mean_diag <- round(median(diag_tb$value[diag_tb$variable=="Diagnosed"]), 2)
lower_diag <- round(quantile(diag_tb$value[diag_tb$variable=="Diagnosed"], .25), 2)
upper_diag <- round(quantile(diag_tb$value[diag_tb$variable=="Diagnosed"], .75), 2)
annon_diag <- paste0("Median diagnosed = ", mean_diag, " (IQR ", lower_diag, " – ", upper_diag, ")")

diag_tb_pl <- diag_tb %>%
  ggplot(aes(x = date, y = value, fill = date_in_model, alpha = variable)) +
  geom_bar(position = position_stack(), stat = "identity") +
  annotate("text", x = as.Date("2021-11-04"), y = 12, label = annon_diag, size = 14/cm(1)) +
  scale_x_date(date_labels = "%b %d", breaks = modeled_days) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c(wes_palette("Moonrise3"), "grey")) +
  scale_alpha_discrete(breaks = c("Diagnosed", "Suspected"), range = c(.33, 1)) +
  labs(y = "Number of patients",
       title = "Daily number of TB patients attending the clinic",
       subtitle = "Diagnosed and suspected (tested) patients in October and November") +
  theme_custom(14) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.direction = "vertical", 
        legend.box = "vertical",
        axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  guides(fill = guide_legend(nrow = 1, order = 2),
         alpha = guide_legend(nrow = 1, order = 1))

save_plot(diag_tb_pl, pdf_file = "../results/data/diagnosed-tb-patients.pdf", w = 18, h = 12)
```

```{r}
diag_tb %>%
  ggplot(aes(x = n_infectious)) +
  geom_histogram()

#' Assumption:
#' Number of diagnosed ~ number of diagnosed TB
#' --> Assume a Dirichlet prior with support according to the counts in the range (1,5).

table(diag_tb$n_infectious)

source("../utils/trans_risk.r")

undiag_tb_pl <- data.frame(x = rundiag(1e5)) %>%
  group_by(x) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(p = 100 * n / sum(n),
         p = round(p)) %>%
  ggplot(aes(x = factor(x), y = p)) +
  geom_bar(stat = "identity", color = wes_palette("GrandBudapest1")[3], fill = wes_palette("GrandBudapest1")[4], alpha = .5) +
  labs(x = "Number of undiagnosed TB patients", y = "Probability (%)",
       title = "Undiagnosed TB patients among clinical attendees",
       subtitle = "Assuming that no. undiagnosed similar to no. diagnosed",
       caption = "Note: Distribution derived from empirical clinical data") +
  coord_cartesian(expand = F) +
  scale_y_continuous(limits = c(0,NA)) +
  theme_custom(14) +
  theme(plot.caption = element_text(size = 10))

undiag_tb_pl
save_plot(undiag_tb_pl, pdf_file = "../results/data/undiagnosed-tb-patients.pdf", w = 14, h = 12)
```