### 2024-03-04 illustrations/priors.Rmd

---
title: "Priors"
author: "Nicolas Banholzer"
date: "2023-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(LaplacesDemon)
source("utils/trans_risk.r")
source("utils/plotting.r")
```

## Generation rate

```{r}
set.seed(12345)

q_dat <- rbind(
  data.frame(
    x = rq(bact_load.mtb(1e6), activity = "waiting"),
    activity = "Waiting"
  ),
  data.frame(
    x = rq(bact_load.mtb(1e6), activity = "walking"),
    activity = "Walking"
  )
)

q_dat %>%
  group_by(activity) %>%
  tidybayes::median_qi(x, .width = .5)

q_dat_dens <- q_dat %>%
  group_by(activity) %>%
  summarize(dens = list(density(x, adjust = 2)))

prev_est <- data.frame(
  study = c(
    "Andrews (2014)",
    "Escombe (2008)",
    "Nardell (1991)",
    "Riley (1962)"
  ),
  x = c(0.89, 8.2, 12.5, 1.25),
  y = c(1, 7, 9, 3) / 1e2
) %>%
  mutate(lab = paste0(study, ": ", round(x, 2)))

q_pl <- ggplot() +
  geom_density(
    data = q_dat,
    mapping = aes(x = x, linetype = activity),
    adjust = 2
  ) +
  geom_segment(
    data = prev_est,
    mapping = aes(x = x, xend = x, y = 0, yend = y, color = study),
    linetype = "dotted"
  ) +
  geom_point(
    data = prev_est,
    mapping = aes(x = x, color = study, y = y),
    shape = 17, size = 2
  ) +
  geom_text(
    data = prev_est,
    mapping = aes(x = x, label = lab, color = study, y = y),
    size = 8 / cm(1), vjust = -2, hjust = 1
  ) +
  scale_x_log10(
    expand = expansion(mult = c(0, .025)),
    breaks = c(1e-4, 1e-5, 1e-3, 1e-2, 1e-1, 1e0, 1e1, 1e2, 1e3, 1e4, 1e5)
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    labels = scales::percent_format(suffix = "")
  ) +
  scale_color_manual(values = wes_palette("Cavalcanti1")) +
  labs(
    x = "Quanta generation rate (quanta/h)",
    y = "Density"
  ) +
  theme_custom() +
  theme(legend.position = "none")

save_plot(
  q_pl,
  pdf_file = "results/inputs/quanta-generation-rate.png",
  w = 16, h = 12
)
```

## Inactivation rate

```{r}
set.seed(12345)

lambda_dat <- data.frame(x = rlambda(1e6))

lambda_dat %>%
  tidybayes::median_qi(x, .width = .5)

prev_est_vir <- data.frame(
  study = c(
    "Loudon (1969)",
    "Lever (2000)",
    "Gannon (2007)",
    "Klein (2014)"
  ),
  x = c(
    log(2) / 6,
    log(2) / (5 / 60),
    log(2) / 1.5,
    log(2) / (0.5 / (log(20 / 100, 0.5)))
  ),
  y = c(15, 5, 10, 20) / 1e2
) %>%
  mutate(lab = paste0(study, ": ", round(x, 2)))

lambda_pl <- ggplot() +
  geom_density(
    data = lambda_dat,
    mapping = aes(x = x),
    adjust = 2
  ) +
  geom_segment(
    data = prev_est_vir,
    mapping = aes(x = x, xend = x, y = 0, yend = y, color = study),
    linetype = "dotted"
  ) +
  geom_point(
    data = prev_est_vir,
    mapping = aes(x = x, color = study, y = y),
    shape = 17, size = 2
  ) +
  geom_text(
    data = prev_est_vir,
    mapping = aes(x = x, label = lab, color = study, y = y),
    size = 8 / cm(1), vjust = -2
  ) +
  scale_x_log10(
    expand = expansion(mult = c(0, .025)),
    breaks = c(1e-4, 1e-5, 1e-3, 1e-2, 1e-1, 1e0, 1e1, 1e2, 1e3, 1e4, 1e5)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, .05)),
    labels = scales::percent_format(suffix = "")
  ) +
  scale_color_manual(values = wes_palette("Cavalcanti1")) +
  labs(
    x = "Bacterial inactivation rate (quanta/h)",
    y = "Density"
  ) +
  theme_custom() +
  theme(legend.position = "none")

save_plot(
  lambda_pl,
  pdf_file = "results/inputs/bacterial-inactivation-rate.png",
  w = 16, h = 12
)
```