### 2024-03-04 analysis/patient-tracking.Rmd

---
title: "Patient movements"
author: "Nicolas Banholzer"
date: "2023-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(lubridate)
library(parallel)
library(reshape2)

source("utils/plotting.r")
source("utils/spatial.r")
```

## Data

### Tracking

```{r}
df <- readRDS("data-clean/patient-tracking/augmented-and-combined-data.rds")

# filter considered clinical attendees
df_matched <- do.call(
  rbind,
  lapply(
    unique(df$date),
    function(d) {
      readRDS(paste0(
        "data-clean/patient-tracking/matched-clinical/",
        d, ".rds"
      ))
    }
  )
) %>%
  mutate(date = as.Date(time))
df_filt <- filter(
  df,
  paste(date, patient_id) %in%
    unique(paste(df_matched$date, df_matched$patient_id))
)
```

### Building

```{r}
cellSize <- 250
source("preprocessing/prep_building-rasterize.R")
```

## Descriptives

```{r}
n_distinct(paste(df$date, df$patient_id))
n_distinct(paste(df_filt$date, df_filt$patient_id))
```

### Tracking quality

```{r}
track_qual <- df %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(tracking_end) %>%
  unlist() %>%
  table()

sum(track_qual[names(track_qual) != "Possible HCW"])
track_qual_str <- paste0(
  names(track_qual),
  ": ",
  track_qual,
  " (",
  round(track_qual / sum(track_qual) * 100),
  "%)"
)
track_qual_str

df %>%
  filter(tracking_end != "Possible HCW") %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(tracking_end) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100))
```

### Time in clinic

```{r}
time_in_clin <- df_filt %>%
  group_by(date, patient_id) %>%
  summarize(
    tracking_end = first(tracking_end),
    `Clinic` = n(),
    `Waiting room` = sum(is_waitingroom),
    `TB room` = sum(is_tbroom),
    `Corridor` = sum(is_passage)
  ) %>%
  ungroup() %>%
  mutate(across(c(`Clinic`, `Waiting room`, `TB room`, `Corridor`), ~ .x / 60))

time_in_clin %>%
  filter(tracking_end != "Possible HCW") %>%
  tidybayes::median_qi(`Clinic`, .width = .5) %>%
  mutate_if(is.numeric, round)

time_in_clin_pl <- time_in_clin %>%
  filter(tracking_end != "Possible HCW") %>%
  dplyr::select(-tracking_end) %>%
  melt(c("date", "patient_id")) %>%
  mutate(variable = factor(
    variable,
    levels = c("Clinic", "Waiting room", "Corridor", "TB room")
  )) %>%
  ggplot(aes(x = variable, fill = variable, y = value)) +
  geom_boxplot() +
  scale_y_sqrt(breaks = c(0, 1, 10, 30, 60, 120, 180)) +
  scale_fill_manual(values = wes_palette("Royal2")) +
  labs(
    y = "Time in clinic (min)",
    title = "c"
  ) +
  theme_custom(8) +
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title.position = "plot"
  )

save_plot(
  time_in_clin_pl,
  pdf_file = "results/data/time-in-clinic.png", w = 5.3, h = 6
)
```

### Close contacts

```{r}
# number of close contacts <1m for >1min
no_close_contacts <- df_filt %>%
  unnest(cols = closeContacts) %>%
  group_by(date, patient_id, closeContacts, tracking_end) %>%
  summarise(
    n_time = n(),
    tracking_end = first(tracking_end)
  ) %>%
  ungroup() %>%
  group_by(date, patient_id) %>%
  summarize(
    n_cc = sum(n_time > 60),
    tracking_end = first(tracking_end)
  ) %>%
  ungroup()

no_close_contacts %>%
  filter(tracking_end != "Possible HCW") %>%
  tidybayes::median_qi(n_cc, .width = .5) %>%
  mutate_if(is.numeric, round)

time_close_contacts <- df_filt %>%
  mutate(has_closeContact = ifelse(
    sapply(closeContacts, length) >= 1, T, F
  )) %>%
  group_by(date, patient_id) %>%
  summarize(
    cc_time = sum(has_closeContact) / 60,
    tracking_end = first(tracking_end)
  ) %>%
  ungroup()

close_contact <- no_close_contacts %>%
  left_join(time_close_contacts) %>%
  left_join(time_in_clin)

cc_n_pl <- close_contact %>%
  filter(tracking_end != "Possible HCW") %>%
  ggplot(aes(x = n_cc)) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    y = "Density", x = "Number of close contacts",
    title = "d"
    # subtitle = "Defined as <1m for >1min"
  ) +
  geom_density(fill = "grey70", color = "grey0", alpha = 0.8) +
  theme_custom(8) +
  theme(
    plot.title.position = "plot",
    plot.margin = margin(r = 11),
    plot.subtitle = element_text(margin = margin(b = 11))
  )

save_plot(
  cc_n_pl,
  pdf_file = "results/data/number-close-contacts.png", w = 5.3, h = 6
)

close_contact %>%
  filter(tracking_end != "Possible HCW") %>%
  tidybayes::median_qi(cc_time / Clinic * 100, .width = .5) %>%
  mutate_if(is.numeric, round)

cc_t_pl <- close_contact %>%
  filter(tracking_end != "Possible HCW") %>%
  ggplot(aes(x = cc_time / Clinic * 100)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    y = "Density",
    x = "Time in close contact (%)",
    title = "e"
  ) +
  geom_density(
    fill = "grey70", color = "grey0", alpha = 0.8
  ) +
  theme_custom(8) +
  theme(
    plot.title.position = "plot",
    plot.margin = margin(r = 11)
  )

save_plot(
  cc_t_pl,
  pdf_file = "results/data/time-close-contacts.png", w = 5.3, h = 6
)
```

### Temporal hotspots

```{r}
clinic_occupancy <- df %>%
  group_by(date, time) %>%
  summarize(
    `Clinic` = n(),
    `Waiting room` = sum(is_waitingroom),
    `TB room` = sum(is_tbroom),
    `Corridor` = sum(is_passage)
  ) %>%
  ungroup() %>%
  melt(c("date", "time")) %>%
  mutate(
    variable = factor(
      variable,
      levels = c("Clinic", "Waiting room", "TB room", "Corridor")
    ),
    date = format(date, "%b %d"),
    date = factor(
      as.character(date),
      levels = c("Oct 13", "Oct 15", "Oct 25", "Nov 04", "Nov 05")
    ),
    time = as.POSIXct(format(time, format = "%H:%M:%S"), format = "%H:%M:%S")
  ) %>%
  filter(between(hour(time), 7, 16))

time_pl <- clinic_occupancy %>%
  filter(variable == "Clinic") %>%
  ggplot(aes(x = time, y = value, color = date)) +
  geom_step() +
  # ggh4x::facet_wrap2(~variable, axes = "x", remove_labels = "all") +
  coord_cartesian(expand = FALSE) +
  scale_x_datetime(date_breaks = "3 hours", date_labels = "%H:%M") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  labs(
    y = "Number of tracks",
    title = "b"
  ) +
  theme_custom(8) +
  theme(
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.key.width = unit(0.275, "cm"),
    legend.key.height = unit(0.2, "cm")
  )

save_plot(
  time_pl,
  pdf_file = "results/data/no-people-over-time.png",
  w = 8, h = 6
)
```

### Spatial hotspots

```{r}
spatial_pl <- plot_spatial(
  ggplot(data = df, mapping = aes(x = x, y = y)) +
    geom_bin2d(binwidth = 200),
  text_descr = 8
) +
  labs(
    fill = "Number of tracks",
    title = "a"
  ) +
  scale_fill_gradientn(
    colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"),
    trans = "log10",
    labels = scales::comma,
    breaks = c(1, 10, 100, 1000, 10000)
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(0.2, "cm"),
    legend.key.width = unit(1, "cm")
  ) +
  guides(fill = guide_colorbar(title.vjust = 1))

save_plot(
  spatial_pl,
  pdf_file = "results/data/no-people-spatial.png",
  w = 8, h = 6
)
```

### Example tracks

```{r}
# unique id
df$day_pid <- paste(df$date, df$patient_id)

# random sample tracks
ex_pid <- df %>%
  filter(tracking_end != "Possible HCW") %>%
  group_by(day_pid) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(tracking_end) %>%
  sample_n(3) %>%
  ungroup() %>%
  dplyr::select(day_pid) %>%
  unlist()
ex_df <- df %>% filter(day_pid %in% ex_pid)
ex_df_se <- ex_df %>%
  group_by(day_pid) %>%
  slice(1, n()) %>%
  mutate(type = factor(c("Start", "End"), levels = c("Start", "End"))) %>%
  ungroup()

# manual sample of tracks
ex_pid_man <- c(
  "2021-10-25 4649", # Entered + Exited, including TB room stay
  "2021-10-13 6297", # Lost Exit, only waiting room
  "2021-11-04 12860", # (Entered) + Exited, waiting room and corridor
  "2021-10-13 9346", # Entered + Exited, including TB room
  "2021-11-05 16625", # Entered + (Exited), only waiting room
  "2021-10-25 9526", # Lost Exit, waiting room and corridor
  "2021-11-04 9841", # Entered = (Exited), waiting room and corridor
  "2021-10-15 8644", # Entered + Exited, including TB room
  "2021-10-15 9419" # Lost Enter
)
ex_df_man <- df %>% filter(day_pid %in% ex_pid_man)
ex_df_man_se <- ex_df_man %>%
  group_by(day_pid) %>%
  slice(1, n()) %>%
  mutate(type = factor(c("Start", "End"), levels = c("Start", "End"))) %>%
  ungroup()

# plot
ex_tracks <- plot_spatial(
  ggplot() +
    geom_path(
      data = ex_df_man,
      mapping = aes(x = x, y = y),
      color = "darkred"
    ) +
    geom_point(
      data = ex_df_man_se,
      mapping = aes(x = x, y = y, shape = type),
      size = 2
    ) +
    facet_wrap(~day_pid, ncol = 3) +
    scale_shape_manual(values = c(1, 13)),
  text_descr = 8
) +
  theme(
    legend.position = "none"
  )

save_plot(
  ex_tracks,
  pdf_file = "results/data/example-patient-tracks.png",
  w = 16, h = 12
)
```