### 2024-03-21 analysis/qctr.Rmd
Results and prior vis updates

---
title: "Quanta concentration and Transmission risk"
author: "Nicolas Banholzer"
date: "2023-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(reshape2)
library(rstanarm)
library(tidybayes)
library(gganimate)

source("utils/plotting.r")
source("utils/tex.r")
source("models/stm-v2.R")

# building
cellSize <- 250
source("preprocessing/prep_building-rasterize.R")

# load tracking and environmental data
track <- readRDS("data-clean/patient-tracking/augmented-and-combined-data.rds")
aer <- readRDS("data-clean/environmental/air-change-rate.rds") %>%
  mutate(daytime = factor(daytime, levels = c("Morning", "Afternoon")))
```

## Illustration

```{r illustration}
# data
mth <- 10
dy <- 25
sim <- 1
param_1025 <- readRDS(
  paste0(
    "../simulations/", "/", "sim-param-",
    2021, "-", mth, "-", dy, ".rds"
  )
)
track_1025 <- readRDS(
  paste0(
    "../simulations/", "/", "tracking-df-",
    250, "px-", 2021, "-", mth, "-", dy, ".rds"
  )
)
track_undiag <- track_1025 %>%
  dplyr::filter(tracking_end != "Possible HCW")
track_diag <- dplyr::filter(track_1025, is_infectious)
track_undiag <- dplyr::filter(
  track_undiag, !is_infectious | is.na(is_infectious)
)
track_tb_all <- rbind(
  track_diag,
  filter(
    track_undiag,
    patient_id %in% param_1025[[paste0("undiagTB", "_", "du")]][[sim]]
  )
)
aer_df <- readRDS("../data-clean/environmental/air-change-rate.rds") %>%
  filter(
    month(date) == as.numeric(mth),
    day(date) == as.numeric(dy)
  )

# time
date <- paste0("2021-", mth, "-", dy)
t_start <- as.POSIXct(paste(date, "08:00:00"))
midday <- as.POSIXct(paste(date, "12:00:00"))
t_end <- as.POSIXct(paste(date, "16:00:00"))
tt <- seq.POSIXt(t_start, t_end, by = "1 sec")
tt_morn <- seq.POSIXt(t_start, midday, by = "1 sec")
tt_noon <- seq.POSIXt(midday + 1, t_end, by = "1 sec")
midday_idx <- which(tt == midday)

# simulation
ct_by_room <- list()
for (room in c("Waiting room", "Corridor", "TB room")) {
  print(paste0("- ", room))

  # room setup
  if (room == "Waiting room") {
    # initial quanta concentration matrix
    c0 <- sP_to_matrix(waiting_room)

    # (no) TB patients in room
    track_tb <- filter(track_tb_all, is_waitingroom)

    # room dimension
    roomDim <- dimWR
  } else if (room == "Corridor") {
    # initial quanta concentration matrix
    c0 <- sP_to_matrix(corridor)

    # TB patients in room
    track_tb <- filter(track_tb_all, is_passage)

    # room dimension
    roomDim <- dimCD
  } else if (room == "TB room") {
    # initial quanta concentration matrix
    c0 <- sP_to_matrix(tb_room)

    # TB patients in room
    track_tb <- filter(track_tb_all, is_tbroom)

    # room dimension
    roomDim <- dimTB
  } else {
    stop("Error: Room invalid.")
  }

  # add quanta generation rate per second by activity level
  track_tb <- track_tb %>%
    dplyr::select(time, cell_x, cell_y, activity) %>%
    rename(x = cell_x, y = cell_y) %>%
    mutate(
      q = ifelse(activity == 1,
        param_1025$q_wait[param_1025$sim == sim],
        param_1025$q_walk[param_1025$sim == sim]
      ),
      q = q * param_1025$mask_red[param_1025$sim == sim],
      q = q / 3600
    )

  # quanta conc in the morning
  print("-- morning")
  track_tb_morn <- filter(track_tb, time <= midday)
  track_tb_morn$t <- as.numeric(difftime(
    track_tb_morn$time, t_start,
    units = "sec"
  ))

  ct_by_room[[room]] <- stm(
    time = tt_morn,
    c0 = c0,
    inf = dplyr::select(track_tb_morn, -time, -activity),
    cellLength = convert_dist(cellSize),
    pd = 0,
    vol = prod(roomDim),
    aer = aer_df$aer[aer_df$room == room & aer_df$daytime == "Morning"] / 3600,
    lambda = param_1025$viral_inact[param_1025$sim == sim] / 3600,
    deposit = 0
  )
}

# get subset
subset_qc <- function(t) {
  # quanta conc
  qc_wr <- melt(ct_by_room[["Waiting room"]][, , t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "Waiting room", across(c(x, y), as.numeric))
  qc_cr <- melt(ct_by_room[["Corridor"]][, , t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "Corridor", across(c(x, y), as.numeric))
  qc_tb <- melt(ct_by_room[["TB room"]][, , t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "TB room", across(c(x, y), as.numeric))

  # link to building
  qc <- rbind(qc_wr, qc_cr, qc_tb)
  qc <- left_join(roomCoord, qc, by = c("x", "y", "room")) %>%
    mutate(
      id = paste0("ID", id),
      t = t
    )

  return(qc)
}

track_tb_all %>%
  filter(between(
    time,
    as.POSIXct(paste(date, "09:41:00")),
    as.POSIXct(paste(date, "09:45:00"))
  )) %>%
  ggplot(aes(x = x, y = y, color = factor(patient_id))) +
  geom_path()

tt_selected <- seq.POSIXt(
  as.POSIXct(paste(date, "09:41:00")),
  as.POSIXct(paste(date, "09:45:00")),
  by = "1 sec"
)
tt_selected_idx <- as.numeric(tt_selected - t_start, unit = "secs")
names(tt_selected) <- tt_selected_idx

qc_sub <- do.call(
  rbind,
  parallel::mclapply(tt_selected_idx, subset_qc, mc.cores = 4)
) %>%
  mutate(t = t - min(tt_selected_idx))

# animated plot
qc_anim_pl <- ggplot(
  data = left_join(
    fortify(waiting_room),
    filter(qc_sub, room == "Waiting room"),
    by = c("id")
  ),
  mapping = aes(x = long, y = lat, group = group, fill = value)
) +
  geom_polygon() +
  coord_cartesian(expand = FALSE) +
  scale_fill_gradientn(
    colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"),
    labels = function(x) round(1e3 * x / (0.25^2 * 3), 1)
  ) +
  theme_custom(14) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    panel.background = element_rect(
      fill = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[1]
    ),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) +
  transition_time(t) +
  labs(
    subtitle = "Time (secs): {frame_time}",
    title = "Quanta concentration in the waiting room on Oct 25 between 9:41 and 09:45am", # nolint
    fill = expression("Quanta/m"^3 * " x 10e"^-3)
  )

animate(
  qc_anim_pl,
  duration = 30,
  width = 1.850877 * 600,
  height = 600,
  renderer = gifski_renderer(),
  nframes = length(tt_selected_idx)
)

anim_save("results/modeling/illustrated-quanta-conc-over-time.gif")
```

## Main

```{r main-files}
main_files <- list.files("simulations/main/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
```

### Quanta conc

```{r}
main_qc_files <- main_files[grepl("quanta", main_files)]

mean_qc <- function(files, room, daytime) {
  room_files <- files[grepl(room, files)]
  room_dt_files <- room_files[grepl(daytime, room_files)]
  qc <- simplify2array(lapply(room_dt_files, readRDS))
  mqc <- apply(qc, c(1, 2), mean)
  mqc_df <- melt(mqc) %>%
    rename(
      y = Var1,
      x = Var2
    ) %>%
    mutate(
      room = room,
      daytime = daytime,
      across(c(x, y), as.numeric)
    )
}

# total quanta concentration by daytime averaged across simulations
main_qc_wr_mo <- mean_qc(main_qc_files, "Waiting room", "morning")
main_qc_wr_af <- mean_qc(main_qc_files, "Waiting room", "afternoon")
main_qc_cr_mo <- mean_qc(main_qc_files, "Corridor", "morning")
main_qc_cr_af <- mean_qc(main_qc_files, "Corridor", "afternoon")
main_qc_tb_mo <- mean_qc(main_qc_files, "TB room", "morning")
main_qc_tb_af <- mean_qc(main_qc_files, "TB room", "afternoon")

# example translation
source("utils/trans_risk.r")
IR_sit <- (.5 * IR("sitting", "male") + .5 * IR("sitting", "female")) / 3600
WR_vol <- convert_dist(cellSize)^2 * dimWR[3]
round((1 - exp(-max(main_qc_wr_mo$value) * 600 * IR_sit / WR_vol)) * 100, 3)
round((1 - exp(-max(main_qc_wr_af$value) * 600 * IR_sit / WR_vol)) * 100, 3)

# link to building
main_qc <- rbind(
  main_qc_wr_mo, main_qc_wr_af,
  main_qc_cr_mo, main_qc_cr_af,
  main_qc_tb_mo, main_qc_tb_af
)
main_qc <- left_join(roomCoord,
  main_qc,
  by = c("x", "y", "room")
) %>%
  mutate(
    id = paste0("ID", id),
    daytime = factor(
      ifelse(daytime == "morning", "Morning", "Afternoon"),
      levels = c("Morning", "Afternoon")
    )
  )

# plot
mean_quanta_pl <- plot_spatial(
  ggplot(mapping = aes(x = long, y = lat, group = group, fill = value)) +
    facet_wrap(~daytime) +
    geom_polygon(
      data = left_join(
        fortify(waiting_room),
        filter(main_qc, room == "Waiting room"),
        by = c("id")
      )
    ) +
    geom_polygon(
      data = left_join(
        fortify(corridor),
        filter(main_qc, room == "Corridor"),
        by = c("id")
      )
    ) +
    geom_polygon(
      data = left_join(fortify(tb_room),
        filter(main_qc, room == "TB room"),
        by = c("id")
      )
    ) +
    labs(
      fill = expression("Mean quanta concentration (quanta m"^-3 * " " * "x 10e"^-3 * ")"),
      title = "a"
    ) +
    scale_fill_gradientn(
      colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"),
      labels = function(x) round(1e3 * x / (0.25^2 * 3), 1)
    ),
  8
) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(0.2, "cm"),
    legend.key.width = unit(2, "cm")
  ) +
  guides(fill = guide_colorbar(title.position = "top"))

mean_quanta_pl
```

### Trans risk

```{r}
average_trans_risk <- function(files) {
  # files
  tr_files <- files[grepl("risk", files)]

  # data
  tr <- do.call(rbind, lapply(tr_files, function(f) {
    readRDS(f) %>% mutate(
      mthdy = stringi::stri_extract(f, regex = "\\d{2}-\\d{2}"),
      sim = stringi::stri_extract(f, regex = "\\d{1,6}")
    )
  }))

  average_tr <- tr %>%
    mutate(patient_id = paste(mthdy, patient_id, sep = "-")) %>%
    group_by(patient_id) %>%
    summarize(
      P_mean = mean(P),
      P_med = median(P)
    ) %>%
    ungroup() %>%
    left_join(track %>%
      mutate(
        mth = lubridate::month(time),
        dy = lubridate::day(time)
      ) %>%
      group_by(patient_id, mth, dy) %>%
      slice(1) %>%
      ungroup() %>%
      mutate(dy = ifelse(nchar(dy) == 1, paste0("0", dy), dy)) %>%
      mutate(patient_id = paste(
        mth, dy,
        patient_id,
        sep = "-"
      )) %>%
      dplyr::select(patient_id, tracking_end)) %>%
    filter(tracking_end != "Possible HCW") %>%
    dplyr::select(-tracking_end) %>%
    reshape2::melt("patient_id") %>%
    mutate(
      val_col = ifelse(value < .001, "Low",
        ifelse(value < .01, "Medium", "High")
      ),
      val_col = factor(val_col, levels = c("Low", "Medium", "High")),
      val_ann = 1 - (1 - value)^12,
      variable = ifelse(
        grepl("mean", variable),
        "Mean risk of infection",
        "Medium risk of infection"
      ),
      variable = factor(
        variable,
        levels = paste(c("Mean", "Medium"), "risk of infection")
      )
    )

  return(average_tr)
}

# average risk
main_tr <- average_trans_risk(main_files)

main_tr %>%
  filter(variable == "Mean risk of infection") %>%
  group_by(val_col) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 0))

# annotation
annotate_average_trans_risk <- function(tr) {
  roi_annon <- tr %>%
    group_by(variable) %>%
    summarize(
      mean_P = mean(value),
      lower_P = quantile(value, .025),
      upper_P = quantile(value, .975),
      mean_P_ann = mean(val_ann),
      lower_P_ann = quantile(val_ann, .025),
      upper_P_ann = quantile(val_ann, .975)
    ) %>%
    ungroup() %>%
    mutate(
      across(c(mean_P, lower_P, upper_P), ~ round_k(100 * .x, 3)),
      lab = paste0(
        mean_P, "%",
        " (95%-CrI ", lower_P, "% – ", upper_P, "%)"
      ),
      across(c(mean_P_ann, lower_P_ann, upper_P_ann), ~ round_k(100 * .x, 2)),
      lab_ann = paste0(
        mean_P_ann, "%",
        " (95%-CrI ", lower_P_ann, "% – ", upper_P_ann, "%)"
      )
    )

  return(roi_annon)
}

main_annon <- annotate_average_trans_risk(main_tr)

main_annon

# plot
roi_pl <- ggplot(
  data = main_tr %>% filter(variable == "Mean risk of infection"),
  mapping = aes(x = value)
) +
  geom_histogram(
    mapping = aes(fill = val_col),
    binwidth = .0005,
    boundary = 1
  ) +
  geom_text(
    data = main_annon %>% filter(variable == "Mean risk of infection"),
    mapping = aes(x = Inf, y = Inf, label = lab),
    size = 8 / cm(1),
    hjust = 1, vjust = 2
  ) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(2, 3, 1)]) +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(
    labels = function(x) x * 100
  ) +
  scale_y_continuous(
    labels = scales::comma
  ) +
  labs(
    x = "Risk of infection (%)",
    y = "Number of clinical attendees",
    title = "b"
  ) +
  theme_custom() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.key.height = unit(.6, "cm"),
    legend.key.width = unit(.4, "cm"),
    panel.spacing = unit(2, "cm")
  )

roi_pl
```

### Association

```{r}
# process tracking data
track_daytime <- track %>%
  mutate(h = lubridate::hour(time)) %>%
  group_by(date, patient_id) %>%
  summarize(
    dur_morning = sum(h < 12),
    dur_afternoon = sum(h >= 12),
    tracking_end = tracking_end[1]
  ) %>%
  ungroup() %>%
  mutate(daytime = ifelse(dur_morning > dur_afternoon, "Morning", "Afternoon"))

time_in_clin <- track %>%
  group_by(date, patient_id) %>%
  summarize(
    `Clinic` = n(),
    `Waiting room` = sum(is_waitingroom),
    `TB room` = sum(is_tbroom),
    `Corridor` = sum(is_passage)
  ) %>%
  ungroup() %>%
  mutate(across(c(`Clinic`, `Waiting room`, `TB room`, `Corridor`), ~ .x / 60))

no_close_contacts <- track %>%
  unnest(cols = closeContacts) %>%
  group_by(date, patient_id, closeContacts, tracking_end) %>%
  summarise(n_time = n()) %>%
  ungroup() %>%
  group_by(date, patient_id) %>%
  summarize(n_cc = sum(n_time > 60)) %>%
  ungroup()

time_close_contacts <- track %>%
  mutate(has_closeContact = ifelse(
    sapply(closeContacts, length) >= 1, TRUE, FALSE
  )) %>%
  group_by(date, patient_id) %>%
  summarize(cc_time = sum(has_closeContact) / 60) %>%
  ungroup()

# merge data
track_ft <- left_join(
  track_daytime,
  left_join(left_join(
    time_in_clin,
    no_close_contacts
  ), time_close_contacts)
) %>%
  mutate(n_cc = ifelse(is.na(n_cc), 0, n_cc)) %>%
  rename(time_in_clinic = `Clinic`) %>%
  dplyr::select(
    date,
    patient_id,
    tracking_end,
    daytime,
    time_in_clinic,
    n_cc,
    cc_time
  )

assoc_df <- track_ft %>%
  # left_join(
  #   aer %>%
  #     filter(room == "Waiting room") %>%
  #     dplyr::select(date, daytime, aer)
  # ) %>%
  mutate(patient_id = paste(
    lubridate::month(date),
    lubridate::day(date),
    patient_id,
    sep = "-"
  )) %>%
  left_join(
    reshape2::dcast(main_tr, patient_id ~ variable) %>%
      dplyr::select(patient_id, `Mean risk of infection`) %>%
      rename(mean_risk = `Mean risk of infection`)
  )

# model
amod <- stan_betareg(mean_risk ~ time_in_clinic + n_cc + cc_time,
  data = assoc_df %>% mutate(
    n_cc = n_cc + 1, cc_time = cc_time + 1 / 60,
    across(c(time_in_clinic, n_cc, cc_time), log2)
  ),
  seed = 12345, cores = 4
)

amod_est <- spread_draws(amod, time_in_clinic, n_cc, cc_time) %>%
  reshape2::melt(c(".chain", ".iteration", ".draw")) %>%
  mutate(value = exp(value))

amod_est %>%
  group_by(variable) %>%
  summarize(
    lower = quantile(value, .025),
    mean = mean(value),
    upper = quantile(value, .975)
  ) %>%
  mutate(across(c(lower, mean, upper), ~ round_k(.x, 2)))

amod_est_pl <- ggplot(amod_est, aes(y = variable, x = value)) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  stat_interval(point_interval = "mean_qi") +
  stat_summary(
    geom = "point",
    fun = mean,
    size = 2,
    shape = 21,
    fill = "white"
  ) +
  labs(
    x = "OR of modeled risk if 2 x predictor",
    title = "c",
    color = "CrI"
  ) +
  scale_y_discrete(
    labels = c(
      # "Air change rate",
      "Time in clinic",
      "No. close contacts",
      "Time close contact"
    )
  ) +
  scale_color_brewer(
    labels = paste0(c(95, 80, 50), "%")
  ) +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom"
  )

amod_est_pl
```

```{r main-figure}
main_fig <- grid.arrange(
  mean_quanta_pl,
  grid.arrange(roi_pl, amod_est_pl, ncol = 2),
  ncol = 1,
  heights = c(8, 8)
)

ggsave(
  main_fig,
  filename = "results/modeling/main-figure.png",
  width = 16 / cm(1),
  height = 16 / cm(1)
)
```

## Additional analyses

```{r additional-analyses}
# files
files_nm <- list.files("simulations/nomasks/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
files_vent <- list.files("simulations/ventilation/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
files_temp <- list.files("simulations/temporal/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
files_prev <- list.files("simulations/prevalence/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
files_susp <- list.files("simulations/suspected/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)

# average risk
nm_tr <- average_trans_risk(files_nm)
vent_tr <- average_trans_risk(files_vent)
temp_tr <- average_trans_risk(files_temp)
prev_tr <- average_trans_risk(files_prev)
susp_tr <- average_trans_risk(files_susp)

tr_comp <- rbind(
  nm_tr %>% mutate(scenario = "No mask wearing"),
  vent_tr %>% mutate(scenario = "Lower ventilation"),
  prev_tr %>% mutate(scenario = "Prevalence-based model"),
  susp_tr %>% mutate(scenario = "Suspected-based model")
) %>%
  mutate(
    scenario = factor(scenario, levels = c(
      "Spatiotemporal model",
      "Temporal model",
      "No mask wearing",
      "Lower ventilation",
      "Prevalence-based model",
      "Suspected-based model"
    ))
  )

# annotation
nm_annon <- annotate_average_trans_risk(nm_tr)
vent_annon <- annotate_average_trans_risk(vent_tr)
temp_annon <- annotate_average_trans_risk(temp_tr)
prev_annon <- annotate_average_trans_risk(prev_tr)
susp_annon <- annotate_average_trans_risk(susp_tr)

annon_comp <- rbind(
  nm_annon %>% mutate(scenario = "No mask wearing"),
  vent_annon %>% mutate(scenario = "Lower ventilation"),
  prev_annon %>% mutate(scenario = "Prevalence-based model"),
  susp_annon %>% mutate(scenario = "Suspected-based model")
) %>%
  mutate(
    scenario = factor(scenario, levels = c(
      "No mask wearing",
      "Lower ventilation",
      "Suspected-based model",
      "Prevalence-based model"
    ))
  )

# plot mean
mean_roi_comp_pl <- ggplot(
  tr_comp %>%
    filter(variable == "Mean risk of infection"),
  aes(x = value)
) +
  geom_histogram(
    mapping = aes(fill = val_col),
    binwidth = .05,
    boundary = 1
  ) +
  geom_text(
    data = annon_comp %>% filter(variable == "Mean risk of infection"),
    mapping = aes(x = Inf, y = Inf, label = lab),
    size = 8 / cm(1),
    hjust = 1, vjust = 2
  ) +
  facet_wrap(~scenario, ncol = 2) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(2, 3, 1)]) +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(
    breaks = c(0, .05, 0.1, seq(0.2, .5, .1)),
    labels = function(x) x * 100
  ) +
  scale_y_continuous(
    labels = scales::comma
  ) +
  labs(
    x = "Risk of infection (%)",
    y = "Number of clinical attendees"
  ) +
  theme_custom() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.key.height = unit(.6, "cm"),
    legend.key.width = unit(.4, "cm"),
    panel.spacing = unit(2, "cm")
  )

save_plot(
  mean_roi_comp_pl,
  "results/modeling/mean-roi-comparison.png",
  w = 16,
  h = 12
)

# comparison with temporal model
mod_comp <- left_join(
  temp_tr %>% rename(`Temporal model` = value) %>% dplyr::select(-val_col),
  main_tr %>% mutate(`Spatiotemporal model` = value) %>% dplyr::select(-val_col)
) %>%
  mutate(
    diff = `Spatiotemporal model` - `Temporal model`
  ) %>%
  mutate(
    val_col = ifelse(diff > .005, "Higher",
      ifelse(diff < -.005, "Lower", "Similar")
    ),
    val_col = factor(val_col, levels = c("Lower", "Similar", "Higher"))
  )

mod_comp %>%
  filter(variable == "Mean risk of infection") %>%
  group_by(val_col) %>%
  summarize(n = n()) %>%
  mutate(prop = round(n / sum(n) * 100, 0))

mod_comp_annon <- mod_comp %>%
  summarise(
    mean_diff = mean(diff),
    lower_diff = quantile(diff, .025),
    upper_diff = quantile(diff, .975)
  ) %>%
  mutate(across(c(mean_diff, lower_diff, upper_diff), ~ round_k(100 * .x, 1)),
    lab = paste0(
      "Average ", mean_diff, "%",
      " (95%-CrI ", lower_diff, "% – ", upper_diff, "%)"
    )
  )

mod_comp_pl <- mod_comp %>%
  ggplot(aes(x = diff)) +
  geom_histogram(
    mapping = aes(fill = val_col),
    binwidth = .005,
    boundary = 1
  ) +
  geom_text(
    data = mod_comp_annon,
    mapping = aes(x = Inf, y = Inf, label = lab),
    size = 8 / cm(1),
    hjust = 1, vjust = 2
  ) +
  labs(
    y = "Number of clinical attendees",
    x = "Difference between spatiotemporal and temporal model (perc. points)"
  ) +
  scale_y_log10(labels = scales::comma) +
  scale_x_continuous(labels = function(x) scales::percent(x, suffix = "")) +
  scale_fill_manual(values = wes_palette("Zissou1")[c(1, 3, 5)]) +
  coord_cartesian(expand = FALSE) +
  theme_custom() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.key.height = unit(.6, "cm"),
    legend.key.width = unit(.4, "cm")
  )

save_plot(
  mod_comp_pl,
  "results/modeling/model-comparison.png",
  w = 12,
  h = 8
)
```