### 2024-05-16 doc/paper/supplements.tex

\PassOptionsToPackage{top=3cm,left=3cm,right=3cm,bottom=3cm}{geometry}
\documentclass[fleqn,11pt]{wlscirep_supp}

\usepackage[]{minitoc}
\mtcsetdepth{secttoc}{3}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\mtcsettitle{secttoc}{}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
%\usepackage[top=3cm,left=3cm,right=3cm,bottom=3cm]{geometry}% by courtesy of Mico
\usepackage{lmodern}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{colortbl}
\usepackage{siunitx}
\sisetup{
  detect-all,
  detect-weight=true,
  detect-family=true,
  mode=text,
%   detect-inline-family=math,
  group-separator={,},
%   group-minimum-digits={3}			
}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{tabu}
\usepackage{authblk}
\usepackage{mathtools}
\usepackage{overpic}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usepackage{multirow}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{bm}
\usepackage{enumerate}
\usepackage[absolute,overlay%,showboxes
                        ]{textpos}
% \usepackage{caption}
\usepackage[font=small,labelfont=bf,justification=justified]{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{placeins}
\usepackage{makecell, booktabs}
\usepackage{eqparbox}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{setspace}
%\usepackage{comment}
\usepackage[resetlabels,labeled]{multibib}
\newcites{Supp}{References}
\usepackage[sort&compress]{cleveref}
\Crefname{appendix}{Supplement}{Supplements}

%\usepackage{mathabx}

% Table formatting packages
\usepackage{dcolumn} % align decimal points in tables
\newcolumntype{d}[1]{D{.}{.}{#1}}
\usepackage{booktabs} 
\usepackage[flushleft]{threeparttable}
\usepackage{siunitx} % align on decimal point in tables
\usepackage{lineno}
\usepackage{etoolbox}

\usepackage{lscape}
\usepackage{longtable}
\usepackage{arydshln}

% math
\usepackage{amsmath,amsfonts,amssymb}
% additional math symbols
\DeclareFontFamily{U}{mathb}{}
\DeclareFontShape{U}{mathb}{m}{n}{
  <-5.5> mathb5
  <5.5-6.5> mathb6
  <6.5-7.5> mathb7
  <7.5-8.5> mathb8
  <8.5-9.5> mathb9
  <9.5-11.5> mathb10
  <11.5-> mathb12
}{}
\DeclareSymbolFont{mathb}{U}{mathb}{m}{n}
\DeclareMathSymbol{\ulsh}{3}{mathb}{"E8}
\DeclareMathSymbol{\ursh}{3}{mathb}{"E9}
\DeclareMathSymbol{\dlsh}{3}{mathb}{"EA}
\DeclareMathSymbol{\drsh}{3}{mathb}{"EB}

%% Patch 'normal' math environments:
\newcommand*\linenomathpatch[1]{%
  \cspreto{#1}{\linenomath}%
  \cspreto{#1*}{\linenomath}%
  \csappto{end#1}{\endlinenomath}%
  \csappto{end#1*}{\endlinenomath}%
}

\linenomathpatch{equation}
\linenomathpatch{gather}
\linenomathpatch{multline}
\linenomathpatch{align}
\linenomathpatch{alignat}
\linenomathpatch{flalign}

\linenumbers

% PLOS formatting
\makeatletter %only needed in preamble
\renewcommand\Large{\@setfontsize\Large{18pt}{18}}
\renewcommand\large{\@setfontsize\large{16pt}{18}}
\makeatother

\addto\captionsenglish{\renewcommand{\figurename}{Figure}}

% \usepackage{xstring}
% \usepackage{etoolbox}
% \usepackage{caption}

% \captionsetup{labelfont=bf,tableposition=top}

% \makeatletter
% \newcommand\formatlabel[1]{%
%     \noexpandarg
%     \IfSubStr{#1}{.}{%
%       \StrBefore{#1}{.}[\firstcaption]%
%       \StrBehind{#1}{.}[\secondcaption]%
%       \textbf{\firstcaption.} \secondcaption}{%
%       #1}%
%       }

% \patchcmd{\@caption}{#3}{\formatlabel{#3}}
% \makeatother

\renewcommand*{\Affilfont}{\normalsize\normalfont}
\renewcommand*{\Authfont}{\normalfont}

% referencing of unnumbered materials and methods
\newcounter{methods}
\renewcommand{\themethods}{Materials and methods}

% Track changes
%\usepackage[markup=underlined]{changes}
\makeatletter
\@namedef{Changes@AuthorColor}{magenta}
\colorlet{Changes@Color}{magenta}
\makeatother

%=====================================================================% Declare

\DeclareSIUnit\eur{\officialeuro}
\DeclareSIUnit\M{M}
\DeclareSIUnit\k{k}

% Widebar symbol
% \DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
% \DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
% \DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
% \DeclareMathAccent{\widebar}{0}{mathx}{"73}

%=====================================================================% New commands (Macros)

% def
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}

% new command
\newcommand{\smallsim}{\smallsym{\mathrel}{\sim}}
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\newcommand{\specialcellc}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\newcommand\ie{i.\,e.\xspace}
\newcommand\eg{e.\,g.\xspace}
\newcommand{\dd}[1][]{\mathrm{d}#1}
\newcommand{\BK}[1]{{\color{orange}{BK: #1}}}
\newcommand{\figletter}[1]{{{\fontfamily{\sfdefault}\selectfont \textbf{#1}}}}
\newcommand\TODO[1]{{\color{red}#1}}  
\newcommand{\FIX}[1]{{\color{darkgreen}#1}}

% renewcommand
\renewcommand\theadfont{\bfseries}
\renewcommand\theadalign{lc}
\renewcommand\cellalign{tl}

\makeatletter

\newbox\@abstract%
\def\abstitle{\textbf{Abstract}}%
\renewenvironment{abstract}{
  \global\setbox\@abstract\vbox\bgroup%
   \noindent
}{%
   \egroup%
}%

\renewcommand*{\Affilfont}{\normalsize\normalfont}
\renewcommand*{\Authfont}{\normalfont}

\addto\captionsenglish{% Replace "english" with the language you use
  \renewcommand{\contentsname}{List of Texts}
}

\def\@maketitle{%
  \newpage
    {\raggedright\fontsize{18pt}{20pt}\selectfont \@title \par}%
    \vskip 0.5em%
    {\large
      \lineskip .5em%
      \begin{tabular}[t]{l}%
        \raggedright \normalsize\mdseries{\@author} %
      \end{tabular}\par}%
      \vskip 1em
%      \raggedright\Large\abstitle\par
%      \vskip 1em
%    {\unvbox\@abstract\par}%
    \par
  \vskip 0.5em
}

\makeatother

\renewcommand{\thesection}{Text \arabic{section}}
\usepackage{titlesec}
%\titleformat{\section}{\normalfont\Large\bfseries}{Text \thesection.~#1}{1em}{}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

\begin{document}
\doublespacing
\nolinenumbers

\title{\LARGE\singlespacing{\textbf{Supplementary Information} \\ \medskip
A novel spatiotemporal model to estimate \emph{Mycobacterium tuberculosis} transmission risks in crowded indoor settings using environmental, clinical, and patient movement data}}

% author list
\author[1]{Nicolas Banholzer}
\author[2]{Keren Middelkoop}
\author[2]{Juane Leukes}
\author[3]{Ernest Weingartner}
\author[1]{Remo Schmutz}
\author[1]{Kathrin Zürcher}
\author[1,4,5]{Matthias Egger}
\author[2]{Robin Wood}
\author[1*]{Lukas Fenner}

\affil[1]{Institute of Social and Preventive Medicine, University of Bern, Bern, Switzerland}
\affil[2]{Desmond Tutu HIV Centre, Department of Medicine, University of Cape Town, Cape Town, South Africa}
\affil[3]{Institute for Sensors and Electronics, University of Applied Sciences and Arts Northwestern Switzerland, Windisch, Switzerland}
\affil[4]{Population Health Sciences, University of Bristol, Bristol, United Kingdom}
\affil[5]{Centre for Infectious Disease Epidemiology and Research, University of Cape Town, Cape Town, South Africa}

\affil[*]{Corresponding author: lukas.fenner@ispm.unibe.ch }

%\begin{abstract}\normalfont
%The supplementary material contains (1)~the detailed method, (2)~the simulation-based study, (3)~further descriptives, and (4)~the results from the sensitivity analysis.
%\end{abstract}

\flushbottom
\maketitle
\thispagestyle{empty}

%\newpage

\sloppy
\raggedbottom

\newpage

\appendix

\tableofcontents

\listoffigures

\listoftables

%\listoffigures
%\listoftables

\newpage

\section{Processing of person-tracking data}\label{sec:setting-and-data}

Person-tracking data required processing because many tracks were interrupted, for example, if the person i)~moved outside the range of the video sensors and then re-entered, ii)~bowed heavily downwards, or iii)~was temporarily hidden behind another person or obstacle. In addition, moving objects such as opening doors may have been recognized as a person, introducing many extremely short tracks (total duration of less than 10\,s and distance less than 2\,m), which were removed.

We reconnected tracks that most likely belonged to the same person. First, we automatically matched tracks that were less than 10\,s and 3\,m apart while moving or less than 60\,s and 0.5\,m apart while sitting. Second, we manually reconnected the remaining tracks if they were likely from the same person. For that purpose, we developed a Shiny app and a matching manual (\url{https://github.com/nbanho/clinic-transmission/wiki/Matching-Manual}) to assist the students performing the matching.

The Shiny app is shown in \Cref{fig:shiny-app}. In the center panel, a floor plan of the clinic is shown in (x,y)-space. Different areas of the clinic are delineated by colors, \eg seating areas are marked in purple and exit areas are marked in blue. Tracks are shown as lines with a open and crossed circles indicating the tracks' start and end, respectively. The currently selected track is shown in black, possible links in colors, and alternatives for the selected possible link in light grey.

Below the floor plan, a table is showing descriptive information about the possible links in relation to the currently selected track: (1)~the patient ID of the selected track (Pid) and the observation ID of possible links (Oid), (2)~the height difference at the end point of Pid and the start point of Oid, (3)~ the standing height (\ie maximum height during the entire track) difference, (4)~the duration of Oid, (5)~the time difference, and (6)~the distance between the end point of Pid and the start of point of Oid. Possible links are sorted in ascending order by the time difference.

The Shiny app's left panel shows descriptive information about the loaded file, the date of the tracking data, the initial number of IDs before any manual linkages, the current number of IDs after considering manual linkages, and the current number of made linkages. It further shows the current Pid with shortcuts to move to the next or previous Pid, or the next or previous Pid at the entrance. Linkages can be done either forward or backward in time. The total duration of the Pid track (time in the clinic) is shown together with the number of connected tracks within Pid. The sliding bars allow filtering of possible links (Oid) according to difference in time, distance, or height, with quick filters to switch between common thresholds. The first dropdown menu selects the Oid for which alternative matches should be shown in the floor plan. The idea is to prevent false linkages when there are more plausible alternatives for the Oid. A connection is made by selecting the Oid in the "Link with IDs" dropdown menu and then clicking "Link IDs". After that, the app updates the floor plan, matching table, and descriptive information. Erroneous link can be revoked by clicking "Unlink last ID", if the last link in time should be revoked, or by clicking "Unlink first ID", if the first link in time should be revoked. If the Pid could not be connected further, its track can be finished by entering a label and clicking "Label/End track". The person-tracking data is then updated with the reconnected track and saved automatically to the file shown at the bottom.

We created the following labels to denote finished tracks: 
\begin{itemize}
    \item \textbf{Clean}: A clean track throughout the clinic with known entrance and exit.
    \item \textbf{Staff}: IDs with moving in- and out of the staff room.
    \item \textbf{Noise}: A short track ($\leq$1\,min) without entrance or exit.
    \item \textbf{Lost enter}: A long track (>1\,min) without entrance.
    \item \textbf{Lost exit}: A long track (>1\,min) without exit.
    \item \textbf{Lost both}: A long track (>1\,min) without entrance and without exit.
\end{itemize}

\begin{figure}[!htpb]
    \centering
    \includegraphics[width=\linewidth]{doc/paper/shiny-tool-screenshot.png}
    \caption{Screenshot of the Shiny app used for manual linking tracking IDs that probably belong together.}
    \label{fig:shiny-app}
\end{figure}

We obtained 3,319 tracks after preprocessing. Further exclusion of short tracks <5\,min resulted in 1,563 tracks used for analysis: 209 (13\%) labeled as "Clean", 125 (8\%) as "Staff", 213 (14\%) as "Lost enter", 246 (16\%) as "Lost exit", and 770 (49\%) as "Lost both". A sample of person tracks are shown in \Cref{fig:tracking-examples}.

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/data/example-patient-tracks.png}
    \caption{Sample of person tracks from our video sensors after preprocessing.}
    \label{fig:tracking-examples}
\end{figure}

\clearpage

\section{Spatiotemporal modeling approach}\label{sec:spattemp-model}

\subsection{The Wells-Riley modeling framework}

We build upon the Wells-Riley model\cite{Riley1961Book}, which is an established epidemiological model to estimate the risk of infection. The model estimates the risk (probability) of airborne indoor transmission $P$ using the following equation: 
\begin{align}
    P = \frac{C}{S} = 1 - \exp \left(-\frac{Ipqt}{Q}\right),
\end{align}
where $C$ is the number of diseased cases, $S$ is the number of susceptible cases, $I$ is the number of infectious people in the indoor space, $p$ is the breathing rate (m$^3$ h$^{-1}$), $q$ is the quantum (infectious dose) generation rate (quanta h$^{-1}$), $t$ is the exposure time (s), and $Q$ is the ventilation rate (m$^3$ h$^{-1}$).

The unknown parameter $q$ is not the actual number of infectious particles in the air. It rather represents the dose of infectious particles that corresponds to a certain probability of infection with a Poisson relationship, \eg one quantum corresponds to $P = 1 - \exp (1) = 63\%$ risk of infection\cite{Rudnick2003IndoorAir}. It is an attempt to consider the stochastic behavior of airborne transmission, since the number of infectious particles that cause an infection is unknown. Moreover, the risk of transmission depends on the environment, characteristics of the pathogen, and characteristics of the individual. For example, the type of respiratory activity (\eg breathing, coughing, or sneezing) determines the site in the respiratory tract where the exhaled particles are generated, which in turn influences the particle size distribution\cite{Wei2016AMJIC}. Smaller particles reach deeper into the lungs of the susceptible person\cite{Wang2021Science}, which increases the probability of infection.

While considering the stochastic behaviour of airborne transmission, the Wells-Riley model makes two simplifying assumptions. First, it assumes a well-mixed airspace, which means that the generated quanta disperse immediately and uniformly in the indoor airspace. In other words, the exposure to infectious quanta is the same regardless of the location of the infectious and susceptible individual. Second, the Wells-Riley model assumes steady-state conditions, which means that the quanta concentration and the outdoor air supply rate are constant over time. In other words, the risk of infection for each susceptible only varies with the duration of exposure (the time spent in the indoor space).

Rudnick and Milton\cite{Rudnick2003IndoorAir} modified the Wells-Riley model to relax the steady-state assumption. They use CO$_2$ as a biomarker for exhaled breath to compute the outdoor air supply rate, which is otherwise difficult to measure. Their modified equation is
\begin{align}
    P = \frac{C}{S} = 1 - \exp \left(-\frac{\bar{f}Iqt}{n}\right),
\end{align}
where $f = \frac{CO^{\text{I}}-CO^{\text{O}}}{CO^{\text{A}}}$ is the fraction of indoor air that is exhaled breath, which is computed based on the CO$_2$ level in indoor air $CO^{\text{I}}$, outdoor air $CO^{\text{O}}$, and exhaled breath $CO^{\text{A}}$ (in parts per million [ppm], respectively).

The extension by Rudnick and Milton still assumes a well-mixed airspace. We relax this assumption by modeling spatial variation in the concentration of infectious quanta. We then compute the risk of infection for susceptible individuals using the Wells-Riley equation considering their cumulative exposure to infectious quanta over space and time as 
\begin{equation}\label{eq:spattemp-P}
    P = \sum_s \sum_t N_{c,t} \cdot \mathbb{I}_{c,t} \cdot p,
\end{equation}
where $N$ is the quanta concentration in the indoor space area $c$ at time $t$ (in quanta/$m^3$), $\mathbb{I}$ is a binary variable indicating whether the susceptible individual was in $c$ at $t$, and $p$ is the breathing rate.

Modeling of the spatiotemporal quanta concentration can be divided into three processes
\begin{itemize}
    \item[\ref{sec:quanta-generation}] \textbf{Quanta generation}: Infectious particles are generated at the infectious individual's location. 
    \item[\ref{sec:quanta-diffusion}] \textbf{Quanta diffusion}: Pathogen-carrying particles spread in the indoor space. 
    \item[\ref{sec:quanta-removal}] \textbf{Quanta removal}: Contaminated indoor air is replaced or pathogens become inactive. 
\end{itemize}
We describe each process in more detail in the following, but first we introduce some notation.

\subsection{Model setup and notation}

We divide the airspace into a 2-dimensional grid with cells $c = 1, \dots, C$, where $(x_c, y_c)$ refer to the center of cell $c$. The grid provides a discrete approximation to the continuously varying quanta concentration in the indoor space, with the number of grid cells $C = n_x \cdot n_y$ determining the spatial resolution. We denote the following variables and modeling parameters:
\begin{itemize}
    \item $N_{c,t}$: quanta concentration in airspace unit $c = 1, \dots, C$ at time $t = 1, \dots T$ (in seconds)
    \item $I_{c,t}$: number of infectious individuals in airspace unit $c$ at time $t$
    \item $q$: quanta generation rate (quanta s$^{-1}$)
    \item $D$: diffusion constant (m$^2$ s$^{-1}$)
    \item $AER$: air change rate (air changes s$^{-1}$)
    \item $V$: volume of the airspace (m$^3$)
    \item $CO^{\text{I}}$: CO$_2$ concentration in indoor air (parts per million, ppm)
    \item $CO^{\text{O}}$: CO$_2$ concentration in outdoor air (ppm)
    \item $G$: average CO$_2$ generation rate per person (L$\cdot$min$^{-1}\cdot$person$^{-1}$)
    \item $n_t$: number of individuals in the airspace at time $t$
    \item $\lambda$: bacterial inactivation rate (quanta s$^{-1}$)
    \item $k$: gravitational settling rate (quanta s$^{-1}$)
\end{itemize}

\subsection{Quanta generation}\label{sec:quanta-generation}

The number of quanta generated by infectious individuals at time $t$ in cell $c$ is 
\begin{align}\label{eq:generation}
    I_{c,t} \cdot q ~.
\end{align}
We assume that the initially generated quanta is confined to the infectious individual's cell, reflecting previous findings that the concentration of pathogen-laden aerosols is typically highest near the infectious source\cite{Wang2021Science,Morawska2021,Vuorinen2020SafSci,Chen2020BuildEnv}. Although coughing and sneezing can expel the aerosols further away from the individual, these activities are much less frequent than breathing, which may thus contribute more infectious particles\cite{Dinkele2022AJRCCM}.

\subsection{Quanta diffusion}\label{sec:quanta-diffusion}

Absent measurements of airflow, we assume that aerosols diffuse uniformly in the indoor air, approaching a well-mixed airspace over time (the steady state). We assume that the quanta diffuse radially in (x,y)-direction towards locations where the concentration is lower. The diffusion equation is given by 
\begin{align}\label{eq:diffusion}
    \frac{\delta N_{c,t}}{\delta t} = D \Delta N_{c,t},
\end{align}
where $\Delta$ is the Laplace operator and $D$ is the diffusion constant. We solve this equation with a standard second-order finite difference model using the Euler method with a 5-point stencil.

The diffusion constant $D$ is unknown. We approximate it using an empirical relationship between the eddy diffusion coefficient $K$ ($K \approx D$) and the outdoor air exchange rate $AER$ for CO$_2$ particles\cite{Cheng2011EnvSciTech,Foat2020BE}
\begin{align}
    K = (0.52 \cdot AER + 8.61\cdot10^{-5}) \cdot V^{\frac{2}{3}},
\end{align}
where $V$ is the volume of the airspace (in m$^3$). We thus assume that pathogen-carrying particles spread with a similar speed than CO$_2$ particles, which likely overestimates the diffusion constant, because CO$_2$ particles are smaller than \emph{Mtb}-carrying particles.

The $AER$ is not directly observed but can be estimated from indoor CO$_2$ levels, room occupancy, and volume. Under steady-state conditions (\ie CO$_2$ levels reaching a steady-state concentration), the $AER$ can be computed as
\begin{align}
    AER = \frac{6\cdot10^4 \cdot n \cdot G}{V\cdot(CO^{\text{I}}-CO^{\text{O}})},
\end{align}
where $n$ is the number of people in the airspace, $G$ is the average CO$_2$ generation rate per person, $CO^{\text{I}}$ is the CO$_2$ concentration in indoor and $CO^{\text{O}}$ is the concentration in outdoor air\cite{Batterman2017IJERPH}. However, in a primary care clinic, room occupancy varies continuously and thus it may be difficult to determine the steady-state CO$_2$ concentration. Therefore, we determine $AER$ using a transient mass balance model of the form\cite{Batterman2017IJERPH}
\begin{align}
    CO_{t+1}^{\text{I}} = \frac{6\cdot10^4 \cdot n_t + G}{Q} \cdot \left(1 - \exp(-Q/V \Delta t)\right) + (CO_t^{\text{I}}-CO^{\text{O}}) \cdot \exp(-Q/V \Delta t) + CO^{\text{O}},
\end{align}
where $Q$ is the outdoor air supply rate (in m$^3$/h) and $AER = Q/V$. We numerically solve the transient mass balance model with the limited memory Broyden–Fletcher–Goldfarb–Shanno algorithm (L-BFGS), a quasi-Newton method, as implemented in the statistical software R\cite{Byrd1995SIAM}. The optimal solution is the outdoor air supply rate $Q$ that minimizes the root mean square error. We also follow recommendations to fit the outdoor CO$_2$ level with reasonable constraints on $CO^{\text{O}} = [300,600]$\cite{Batterman2017IJERPH}. We can compute $AER$ over different time windows, hence the index in $AER_t$.

\subsection{Quanta removal}\label{sec:quanta-removal}

Infectious quanta may be removed from the air through outdoor air exchange, bacterial inactivation, or gravitational settling. We will ignore gravitational settling because \emph{Mtb} is carried in small particles between 1\,$\mu$m and 7\,$\mu$m\cite{Fennelly2020Lancet}, which become airborne immediately or within seconds\cite{Vuorinen2020SafSci}. The removal rate is thus the sum of the outdoor air exchange rate $AER_t$ and the bacterial inactivation rate $\lambda$ for \emph{Mtb}, \ie 
\begin{align}\label{eq:removal}
    AER_t + \lambda ~.
\end{align}

\subsection{Spatiotemporal quanta concentration}

Combining \Cref{eq:generation}, \Cref{eq:diffusion}, and \Cref{eq:removal}, the quanta concentration at time $t$ in the indoor space can be computed as
\begin{align}\label{eq:spattemp-N}
    \underbrace{N_{t}}_{\text{new concn.}} = \underbrace{\left(D \Delta (\underbrace{N_{t-1}}_{\text{prev. concn.}} + \underbrace{I_t \cdot q}_{\text{generation}})\right)}_{\text{diffusion}} \cdot \underbrace{\exp\left(-(AER_t + \lambda)\right)}_{\text{removal}} ~.
\end{align}

\subsection{Illustrative example}\label{sec:example}

We illustrate our modeling approach with an example. Consider a room with a volume $V$=150\,m$^3$ (length = 10\,m, breadth = 5\,m, height = 3\,m), which we rasterize into a grid of $C = 40 \cdot 20 = 800$\,cells with an area of 0.25\,m$^2$. We assume that quanta generation is confined to the cell where the infectious individual is located as well as the first neighbouring cells. We monitor the room from 8\,am to 12\,am and consider one infectious individual generating quanta continuously for one hour between 8\,am and 9\,am in the middle of the room (cell (20,10)), which means that quanta is generated in cells $(x_c = 20\pm1, y_c = 10\pm1)$. We assume a quanta generation rate of $q = 10$\,quanta h$^{-1}$ and a bacterial inactivation rate of $\lambda = 0.5$\,quanta h$^{-1}$. We further assume a constant outdoor air exchange rate of $AER = 1$\,air changes h$^{-1}$, translating into a quanta removal rate of $1 + 0.5 = 1.5$\,quanta h$^{-1}$ and corresponding to a diffusion constant of 0.007\,m$^2$ s$^{-1}$.

\Cref{fig:toy-example} shows the quanta concentration at different time points. The quanta concentration increases over time and is higher near the infectious individual in the room's center. After the individual has left the room, the quanta diffuses fairly quickly and a well-mixed airspace is already at 9:05\,am. Over time, the removal through outdoor air exchange and bacterial inactivation brings the concentration in the entire room back towards zero.

\begin{figure}[!htpb]
    \centering
    \includegraphics{tests/stm_v2-toy_example.png}
    \caption{Spatiotemporal quanta concentration in an indoor space between 8\,am and 10\,am following the generation of quanta by an infectious individual between 8\,am and 9\,am in the room's center.}
    \label{fig:toy-example}
\end{figure}

\clearpage

\section{Simulations}\label{sec:estimation}

\subsection{Setup and setting}\label{prep:building}

We divide the clinic into three spaces: the waiting room (length = 10.55\,m $\times$ breadth = 5.50\,m $\times$ height = 3.00\,m), the corridor (length = 12.45\,m $\times$ breadth = 2.20\,m $\times$ height = 2.50\,m), and the TB room (length = 4.75\,m $\times$ breadth = 3.50\,m $\times$ height = 3.00\,m). We rasterize each room into a grid of cubic cells with a squared area of 0.25\,m$^2$, corresponding to a diagonal length of $\sqrt{0.25^2 + 0.25^2} = 0.35$\,m.

Although the door between the waiting room and corridor was usually not closed, we consider these rooms as separate spaces to facilitate modeling and because we observed differences in room occupancy and time-varying CO$_2$ levels. Nevertheless, we acknowledge the possibility for aerosolized \emph{Mtb} (quanta) to spread from the waiting room to the corridor, and vice versa, which is not considered in our simulations.

We model the spatiotemporal quanta concentration during clinic hours from 8\,am to 4\,pm and compare the modeled concentration by daytime, \ie morning (8\,am to 12\,am) and afternoon (12\,am to 4\,pm). Correspondingly, we estimate the outdoor air exchange rate by daytime, considering potential differences in ventilation conditions between the morning and afternoon. We update the quanta concentration every second from $t = 1, \dots, T$, in line with the frequency of the person-tracking data. We set the quanta concentration at $t=0$ to zero, \ie $N_{c,0} = 0 ~~ \forall c$, assuming that all quanta from the previous day has been removed from the air before the start of the following clinical day.

\subsection{Monte Carlo simulation}

We use Monte Carlo simulation to estimate the risk of infection, considering uncertainty in several modeling parameters. Each simulation consists of the following steps:
\begin{itemize}
    \item[1.] \textbf{Sample uncertain modeling parameters}: Undiagnosed TB patients among clinic attendees, quanta generation rate, the mask reduction rate, and the bacterial inactivation rate.
    \item[2.] \textbf{Compute quanta concentration}: Based on the generation of infectious quanta by diagnosed and undiagnosed TB patients, we compute the quanta concentration in each room per unit space (cell) over time.
    \item[3.] \textbf{Estimate risk of infection}: Based on their spatiotemporal quanta exposure, we estimate the risk of infection for each clinic attendee using the Wells-Riley equation.
\end{itemize}

Many clinic attendees are potential undiagnosed TB patients and prior uncertainty in the quanta generation, mask reduction, and bacterial inactivation rate is large. To reflect this uncertainty, we perform 5,000 Monte Carlo simulations and compute the individual risk of infection as the mean across these simulations.

\subsection{Modeling assumptions}\label{sec:priors}

\subsubsection{Initial spread}

The initial spread, referred to as the propagation distance, is the maximum distance from the source to the end of the exhaled air before dispersion. An analysis of exhalation activities suggest a propagation distance of 0.7\,m\cite{Tang2013PLoSOne}, but larger distances may be reached depending on the activity (\ie breathing, speaking, coughing). Face masks reduce the propagation distance both by blocking (the mask capturing particles) and deflecting (leakage of particles on the sides of the mask) particles\cite{Tang2009RoyalInt,Hui2012PLoSOne,Mansour2013AerosolMed}. Therefore, we assume that mask wearing confines the quanta generation only to the cell where the infectious individual is located. In our setup, this corresponds to a propagation distance of 0.35\,m (the diagonal cell length) with an initial spread to an area of 0.25$^2$\,m$^2$. Without masks, we assume two times that propagation distance, \ie 0.7\,m (the diagonal lengths of two cells)\cite{Tang2013PLoSOne}. Since the direction of propagation is not observed, we assume uniform spread to all neighbouring cells, \ie the generated quanta is distributed evenly among the first-neighbouring cells from where the infectious individual is located.

\subsubsection{Quanta generation rate}

Previous studies estimated the quanta generation rate of \emph{Mtb} using the Wells-Riley model\cite{Andrews2014JID,Riley1962ARRD,Escombe2008PLoSMed,Nardell1991ARRD,Dharmadhikari2012AJRCCM}. Andrews et al.\cite{Andrews2014JID} reported an estimate of 0.89\,(quanta h$^{-1}$) (range from 0.44 to 5.69), close to the early estimate provided by Riley et al.\cite{Riley1962ARRD} with 1.25\,quanta h$^{-1}$. On the other end, Escombe et al.\cite{Escombe2008PLoSMed} arrived at an average estimate of 8.2\,quanta h$^{-1}$, which, however, could be as high as 226\,quanta h$^{-1}$. Lower rates were observed in non-multidrug-resistant (MDR), smear-negative patients and considerably higher ones in MDR, smear-positive TB patients\cite{Escombe2008PLoSMed}. Dharmadhikari et al.\cite{Dharmadhikari2012AJRCCM} also estimated a high average rate of 138\,quanta h$^{-1}$ in MDR TB patients, and Nardell et al.\cite{Nardell1991ARRD} reported an estimate of 13\,quanta h$^{-1}$ for a single index case. The wide range of estimates within and between studies suggest that the emitted quanta varies considerably with the infectiousness of \emph{Mtb} patients\cite{Wurie2016BMJ}. To consider such variation, Mikszewski et al.\cite{Mikszewski2021GF} used the bacterial load distribution in sputum to estimate the quanta generation rate with the approach developed by Buonanno et al.\cite{Buonanno2020EI}. We have recently adopted this approach to derive quanta generation rate distributions for different activity levels in a school setting\cite{Banholzer2024PGPH}. Here, we follow the same approach to derive a quanta generation rate distribution for sitting and walking activities, assuming 80\% breathing and 20\% speaking during both activities. The resulting prior distributions are shown in \Cref{fig:quanta-distribution} along with the estimates from the literature. The quanta generation rate follows a lognormal prior distribution with median 1.08\,quanta h$^{-1}$ (95\%-credible interval [CrI] 0.003\,quanta h$^{-1}$\,$-$\,386\,quanta h$^{-1}$) while sitting and 2.81\,quanta h$^{-1}$ (95\%-CrI 0.008\,quanta h$^{-1}$\,$-$\,987\,quanta h$^{-1}$) while walking. The activity is determined based on whether the the person had moved by more or less than 0.25\,m (spatial distance between two recorded tracks of the same person).

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/quanta-generation-rate.png}
    \caption{Prior distribution (log scale) for the quanta generation rate (quanta h$^{-1}$) by activity level (sitting and walking), along with reported average estimates from the literature\cite{Andrews2014JID,Riley1962ARRD,Escombe2008PLoSMed,Nardell1991ARRD,Dharmadhikari2012AJRCCM}.}
    \label{fig:quanta-distribution}
\end{figure}

\subsubsection{Mask reduction rate}

Face masks reduce the number of viral-laden aerosols that are exhaled into the air\cite{Milton2013PLoSPathogens,Leung2020NatMed} and should thus lower the generation of infectious quanta. McCreesh et al. assumed a reduction by mask-wearing of 75\% (95\%-confidence interval (56\% to 85\%) based on the findings of Dharmadhikari et al.\cite{Dharmadhikari2012AJRCCM}, which we translate into a $\mathrm{Beta}(24.9, 8.3)$ prior distribution for the mask reduction rate. We assume that masks were always correctly worn by all clinic attendees and staff during the study.

\subsubsection{Bacterial inactivation rate}

The bacterial inactivation rate for \emph{Mtb} has not been precisely estimated. Loudon (1969)\cite{Loudon1969AMRRD} report a half-life for aerosolized TB bacilli of 6\,h ($\lambda$ = 0.12). By contrast, Lever (2000)\cite{Lever2000LettersAppliedMicrobio} reported a half-life of just 5\,min ($\lambda$ = 8.3). The difference could be explained by differences in humidity levels during the study\cite{Lever2000LettersAppliedMicrobio}: Loudon (1969) and Lever (2000) studied survival in an environment with relative humidity of about 50\% and 70\%, respectively. Relative humidity in our study is closer to 50\%, thus the estimate by Loudon (1969) may be more suitable for our study. Furthermore, Gannon (2007)\cite{Gannon2007ResVetSci} report a half-life of 1.5\,h ($\lambda$ = 0.46) at relative humidity of over 75\%, yet the estimate is for \emph{Mtb bovis}, and Klein (2014)\cite{Klein2014IJMyco} observe a 80\% reduction within 30\,min ($\lambda$ = 3.2), which is closer to the estimate by Lever (2000). Considering the variation in these estimates, we model $\lambda$ with a $\mathrm{Lognormal}(\mu = \log 1, \sigma = 1)$ distribution, with a median rate of 1\,quanta h$^{-1}$ (95\%-CrI 0.1\,quanta h$^{-1}$\,$-$\,7.1\,quanta h$^{-1}$). The distribution is shown in \Cref{fig:lambda-distribution} along with the estimates from the literature.

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/bacterial-inactivation-rate.png}
    \caption{Prior distribution (log scale) for the bacterial inactivation rate (quanta h$^{-1}$) of \emph{Mtb} along with reported estimates from the literature\cite{Loudon1969AMRRD,Gannon2007ResVetSci,Klein2014IJMyco,Lever2000LettersAppliedMicrobio}.}
    \label{fig:lambda-distribution}
\end{figure}

\subsubsection{Number of undiagnosed TB patients}

Undiagnosed (subclinical) TB patients may go undetected because they do not show symptoms according to the WHO 4-question symptom screen (cough, fever, weight loss, and night sweats)\cite{Berhanu2023CID}, or the diagnosis could not be made at first presentation based on existing tests\cite{Patterson2024PNAS}. Previous studies suggest that the number of undiagnosed (subclinical) TB patients is roughly similar to the number of diagnosed TB patients\cite{Berhanu2023CID,Moyo2022LancetID}. Therefore, we model the number of undiagnosed TB patients with a Multinomial distribution based on the counts of the daily number of diagnosed TB patients from October to November 2021 (\Cref{fig:undiagnosed-distribution}. We consider all clinic attendees as potential undiagnosed TB patients and sample among them with equal probability. We further assume that undiagnosed and diagnosed TB patients have the same quanta generation rate.

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/undiagnosed-tb-patients.png}
    \caption{Counts (\%) of the daily number of diagnosed TB patients visiting the clinic according to clinical records. The number of diagnosed patients is modeled with a Multinomial distribution based on these counts.}
    \label{fig:undiagnosed-distribution}
\end{figure}

\subsubsection{Breathing rate}

Adams (1993)\cite{Adams1993} reports breathing rates for different activity levels. We use the average breathing rate for the following activity levels: sitting with $p = 0.51$\,m$^3$ h$^{-1}$ and walking (slowly at 2.5 mph) with $p = 1.33$\,m$^3$ h$^{-1}$.

\clearpage

\section{Additional results}\label{sec:additional-results}

% \begin{figure}[!htpb]
%     \centering
%     \includegraphics{results/data/diagnosed-tb-patients.png}
%     \caption{Number of diagnosed TB patients (dark bars) and those suspected of TB (light bars) following symptom screening. Modeled study days in color, other clinic days in October and November in grey.}
%     \label{fig:diagnosed-tb-patients}
% \end{figure}

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/data/co2-levels-over-time.png}
    \caption{CO$_2$ levels by clinic area over time for each study day.}
    \label{fig:co2-levels}
\end{figure}

% \begin{figure}
%     \centering
%     \includegraphics{results/modeling/patient-risk-focus-plot.png}
%     \caption{Mean and 80\%-CrI (uncertainty) of the risk of infection per clinic visit. Highlighted in red is a visit from a clinic attendee with moderate risk of infection and low estimated uncertainty.}
%     \label{fig:patient-risk-focus}
% \end{figure}

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/modeling/mean-roi-assumptions.png}
    \caption{Mean risk of infection per clinic attendee if \textbf{(a)}~the number of infectious people in the clinic corresponded to the prevalence of TB in the population, or \textbf{(b)}~the number of infectious people in the clinic corresponded to the number of bacteriologically confirmed TB patients and those suspected of having TB due to respiratory symptoms when attending the clinic. Low: <0.1\%, Medium: 0.1-1\%, High: >1\% risk of infection.}
    \label{fig:assumptions-results}
\end{figure}

\clearpage

\section{Discussion of factors affecting the risk of airborne infection}\label{sec:depth-discussion}

The generation, dispersion, and survival of infectious particles in the air is subject to considerable uncertainty. Therefore, Wells\cite{Wells1955} introduced the concept of a ``quantum'' (infectious dose) to capture the stochastic nature of airborne transmission. A higher concentration of infectious quanta corresponds to a higher probability for an individual to get infected. The probability of infection is modeled with a Poisson relation and the Wells-Riley model considers multiple determinants for the generation and survival of infectious quanta. Nevertheless, transmission events are difficult to predict as the occurrence depends on multiple other factors that can broadly be categorized into (1)~environmental factors, (2)~physicochemical properties of the pathogen, and the (3)~infectiousness of the source and susceptibility of the host.

\subsection{Environmental factors}

Environmental factors are somewhat considered by our model using CO$_2$ as a tracer gas to estimate the air exchange rate, based on which we model quanta diffusion and removal. The combination of clinical and person-tracking data further allows the spatial modeling of quanta generation. This makes our approach unique compared to existing modeling based on the Wells-Riley equation assuming a well-mixed airspace\cite{Riley1961Book,Rudnick2003IndoorAir}. In our spatiotemporal model, we assume a radial diffusion of quanta, which may not reflect the actual airflow at the primary care clinic. The spatial dispersion of infectious particles is often simulated with computational fluid dynamics (CFD) models\cite{Vuorinen2020SafSci,Jung2021InfectChemo,Li2021BuildEnv}. CFD models can simulate a variety of diffusion patterns, which help researchers understanding the influence of airflow. For example, measuring airflow can identify poorly ventilated indoor spaces where infectious particles circulate inside the room rather than getting replaced with fresh outdoor air\cite{Li2021BuildEnv}. However, airflow is rarely considered in modeling because it is difficult to assess in a naturally ventilated space, such as our primary care clinic, where the airflow depends on the outdoor wind direction and which doors and windows are open.

\subsection{Pathogen-specific factors}

Survival and dispersion of aerosols are further determined by the physicochemical properties of virus-laden aerosols such as particle size, viral load, and other chemical components\cite{Wang2021Science}. Currently, these stochastic properties are only indirectly incorporated in the Wells-Riley modeling framework via the quanta generation rate parameter\cite{Riley1961Book,Rudnick2003IndoorAir}. It is challenging to consider these properties specifically, especially because they can be modified by environmental factors with often unclear, pathogen-specific effects\cite{Songer1967,Chan2011AdvVir,Fernstrom2013JoP,Cox1995Book,Fernstrom2013JoP,Tang2009Interface}. For example, survival of bacteria in aerosols decreases with temperatures $>$24$^{\circ}$C, but the effects of humidity are unclear\cite{Tang2009Interface}. On the one hand, a comparison between two studies suggested that longer airborne survival of \emph{Mtb} could be attributed to lower relative humidity\cite{Loudon1969AMRRD,Lever2000LettersAppliedMicrobio}. On the other hand, airborne \emph{Mtb} was more likely detected in healthcare facilitates with higher relative humidity\cite{Sornboot2019IJTLD,Matuka2021IJERP}. Ambient carbon dioxide levels may further influence aerostability\cite{Haddrell2024NatCommun}.

\subsection{Patient-specific factors}

Previous studies show considerable variation in the generation of infectious quanta by TB infected patients\cite{Escombe2008PLoSMed,Andrews2014JID}. We model such variation by using a wide prior distribution for the quanta generation rate. Our model could be easily extended to incorporate prior information on the infectiousness of each individual TB patient. In general, only patients with active pulmonary TB are able to produce infectious droplets\cite{Rieder1999}. Infectiousness could be determined based on the most recent sputum test result. Sputum smear-positive patients are more likely to transmit \emph{Mtb} than sputum smear-negative patients\cite{Shaw1954ART,Brindle1993AMRRD,Grzybowski1975BIUT}. HIV status could further inform both infectiousness and susceptibility. On the one hand, HIV-positive patients are more likely to have a negative sputum smear microscopy result than HIV-negative patients, due to the reduced lung cavitation and the fewer \emph{Mtb} bacilli in their sputum\cite{Brindle1993AMRRD,Telzak1997CID}. On the other hand, HIV-positive patients are less prone to control the \emph{Mtb} infection than HIV-negative patients, due to their reduced immune status\cite{Forte1992AIDS,Kwan2011CMR,Shen1988CEI}, and therefore TB disease progresses faster and is often more severe in HIV-positive TB patients compared to HIV-negative ones.

% data and other limitations
% While we combined multiple data sources to allow spatiotemporal modeling of the generation, spread, and removal of infectious quanta, our data has also limitations. First, clinical attendees could not always be tracked perfectly throughout the clinic using our video sensors. We invested considerable effort to manually link individual tracking IDs probably belonging to the same attendee back together (Supplementary Text~\zref{sec:setting-and-data} in \supp), but some measurement error remains. Second, we linked video tracking and clinical patient data using the times clinical attendees were at the registration and the entry times in the clinical database. False linkages could result when a tracking ID only passed by the registration (without registering) or when multiple IDs were at the registration closely before a clinical entry. Third, to facilitate modeling, we considered the waiting room and corridor as separate, although the door between both rooms was usually open (Supplementary Text~\zref{sec:estimation} in \supp). Fifth, our video sensors could not cover the whole area of the entrance and we did not collect data in the care rooms and other places of the clinic. As a result, the time spent in the clinic and the exposure to infectious quanta may be underestimated for some clinical attendees, although we expect the largest exposure to be in the waiting room, which was fully covered.

\clearpage

\bibliography{references.bib}

\end{document}